### 2015-02-04 paper.tex
first commit

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{rldmsubmit,palatino}
\usepackage{graphicx}

\title{Inverse Modular Reinforcement Learning on Human Behavior}

\author{
Shun Zhang\\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{menie482@cs.utexas.edu} \\
\And
Matthew Tong \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Mary Hayhoe \\
Affiliation \\
Address \\
\texttt{email} \\
\And
Dana Ballard \\
Affiliation \\
Address \\
\texttt{email} \\
\\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

\maketitle

\begin{abstract}
The \emph{title} should be a maximum of 100 characters.

The \emph{abstract} should be a maximum of 2000 characters of text,
including spaces (no figure is allowed). You will be asked to copy
this into a text-only box; and it will appear as such in the
conference booklet. Use 11~point type, with a vertical spacing of
12~points.  The word \textbf{Abstract} must be centered, bold, and in
point size 12. Two line spaces precede the abstract.
\end{abstract}

\keywords{
enter key words here
}

\acknowledgements{We are deeply indebted to iCORE for
their generous support of RLDM2015.}

\startmain % to start the main 1-4 pages of the submission.

\section{Introduction}

Consider the task illustrated in Figure~\ref{fig:avatar} A). The avatar is asked to
do three sub-tasks simultaneously --- 1) following the path, indicated by the gray
line on the ground, 2) getting targets, the blue cylinders, and 3) avoiding
obstacles, the pink cylinders \cite{rothkopf2013modular}.

From the reinforcement learning perspective, this task can be decomposed to be
three sub-tasks as described above.  In Figure~\ref{fig:avatar} B), if the agent
knows the distance and angle to an object, he is expected to know the optimal
action to avoid or pursue it.

A human has not done this kind of task before can achieve a good performance
easily, shown in the experiments described later. If a human knows the policies
of the sub-tasks, or sub-MDPs, he can accomplish a complicated behavior by
combining the sub-MDPs. That is,
$$Q(s, a) = \sum_i w_i Q_i (s, a)$$
where $Q_i$ is the Q value of the i-th sub-MDP, $w_i$ is the weight of the i-th
sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.

% scheduling

\begin{figure}[h!]
\includegraphics[width=0.5\textwidth]{task_1_room_2.png}
\includegraphics[width=0.5\textwidth]{task_4_room_25.png}
\caption{}
\label{fig:exp}
\end{figure}

Different weights can yield different performance. Let $w_1, w_2, w_3$ be
weights for the task of target collection, obstacle avoidance, and path
following, respectively. Let $w$ be the vector of $(w_i)_1^n$. An agent with $w
= (1, 0, 0)$ only collect targets, and one with $w = (0, 0.5, 0.5)$ may avoid
the obstacles and follow the path.

From a different perspective, can we find a weight vector to best interpret
human's behavior? In Figure~\ref{fig:exp}, there are two scenarios. Same as
Figure~\ref{fig:avatar}, the red circles are obstacles. The blue circles are
targets. The gray line is the path. The black lines are trajectories of human
data. The left figure shows the case where human follows the path and ignores
the targets and obstacles. The right figure shows the case that human does three
sub-tasks simultaneously. The human data are collected by the Center for
Perceptual Systems in University of Texas at Austin.

Now, we assume a learning agent that only knows the three sub-MDPs and the human
data. It looks at the human behavior, and finds the weights that can interpret
such behavior. Using such weights, the trajectories of our agents are drawn in
the green lines. We can tell that in the left figure, the agent puts a large
weight on path-following. In the right figure, it puts weights on all sub-MDPs.

\bibliographystyle{plain}
\bibliography{paper}

\end{document}

### 2015-02-11 paper.tex
added basic content

@@
 
-\title{Inverse Modular Reinforcement Learning on Human Behavior}
+\title{Inverse Modular Reinforcement Learning on Human Motion}
 
@@
 Matthew Tong \\
-Affiliation \\
-Address \\
-\texttt{email} \\
-\And
+Center for Perceptual Systems\\
+University of Texas at Austin\\
+Austin, TX 78712 \\
+\texttt{mhtong@gmail.com} \\
+\AND
 Mary Hayhoe \\
-Affiliation \\
-Address \\
-\texttt{email} \\
+Center for Perceptual Systems\\
+University of Texas at Austin\\
+Austin, TX 78712 \\
+\texttt{hayhoe@utexas.edu} \\
 \And
 Dana Ballard \\
-Affiliation \\
-Address \\
-\texttt{email} \\
+Department of Computer Science\\
+University of Texas at Austin\\
+Austin, TX 78712 \\
+\texttt{dana@cs.utexas.edu} \\
 \\
@@
 \begin{abstract}
-The \emph{title} should be a maximum of 100 characters.
-
-The \emph{abstract} should be a maximum of 2000 characters of text,
-including spaces (no figure is allowed). You will be asked to copy
-this into a text-only box; and it will appear as such in the
-conference booklet. Use 11~point type, with a vertical spacing of
-12~points.  The word \textbf{Abstract} must be centered, bold, and in
-point size 12. Two line spaces precede the abstract.
+%The \emph{abstract} should be a maximum of 2000 characters of text, including
+%spaces (no figure is allowed).
 \end{abstract}
@@
 
-\acknowledgements{We are deeply indebted to iCORE for
-their generous support of RLDM2015.}
+\acknowledgements{}
 
@@
 \section{Introduction}
+
+Human is able to learn complicated behaviors much faster than machines can do.
+
+We analyze human's behavior of accomplishing various tasks. With the best of our
+knowledge, there is no work using inverse reinforcement learning to understand
+human motional behavior in the literature.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.6\textwidth]{avatar.png}
+\caption{The task of collecting targets, avoiding obstacles and following the
+path.}
+\label{fig:avatar}
+\end{figure}
 
@@
 A human has not done this kind of task before can achieve a good performance
-easily, shown in the experiments described later. If a human knows the policies
+easily, shown in the experiments described later.
+
+\section{Modular Reinforcement Learning}
+
+Value functions is decomposed \cite{koller1999computing}.
+
+ If a human knows the policies
 of the sub-tasks, or sub-MDPs, he can accomplish a complicated behavior by
@@
 sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-
-% scheduling
-
-\begin{figure}[h!]
-\includegraphics[width=0.5\textwidth]{task_1_room_2.png}
-\includegraphics[width=0.5\textwidth]{task_4_room_25.png}
-\caption{}
-\label{fig:exp}
-\end{figure}
 
@@
 
+\section{Experiments}
+
+We conducted experiments to ask volunteers to accomplish different tasks and
+recorded their trajectories. There are four kind of tasks. Task 1, following the
+path only, and ignoring other objects. Task 2, following the path, while avoid
+the obstacles. Task 3, following the path, while attain targets when possible.
+Task 4, following the path, collecting the targets and avoiding obstacles
+simutanously.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.24\textwidth]{task_1.png}
+\includegraphics[width=0.24\textwidth]{task_2.png}
+\includegraphics[width=0.24\textwidth]{task_3.png}
+\includegraphics[width=0.24\textwidth]{task_4.png}
+\caption{}
+\label{fig:exp}
+\end{figure}
+
 From a different perspective, can we find a weight vector to best interpret
-human's behavior? In Figure~\ref{fig:exp}, there are two scenarios. Same as
-Figure~\ref{fig:avatar}, the red circles are obstacles. The blue circles are
-targets. The gray line is the path. The black lines are trajectories of human
-data. The left figure shows the case where human follows the path and ignores
-the targets and obstacles. The right figure shows the case that human does three
-sub-tasks simultaneously. The human data are collected by the Center for
-Perceptual Systems in University of Texas at Austin.
+human's behavior. In Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar}, the
+red circles are obstacles. The blue circles are targets. The gray line is the
+path. The black lines are trajectories of human . The left figure shows the case
+where human follows the path and ignores the targets and obstacles. The right
+figure shows the case that human does three sub-tasks simultaneously. The human
+data are collected by the Center for Perceptual Systems in University of Texas
+at Austin.
+
+We make some constraints on our learning agent to make it behavior like humans.
+It is allowed to go straight ahead, turn left with a small step, or turn right
+with a small step.
 
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
+\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
+\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
+\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
+\caption{Heatmaps of the objective values of different weights for the four
+tasks, respectively. The red zones indicate higher values. The upper two are
+Task 1 and 2. The lower two are Task 3 and 4.}
+\label{fig:heatmap}
+\end{figure}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{| l| l| l |}
+\hline
+Average by Task & Num Targs Hit & Num Obst Hit\\
+\hline
+1 & 2.34 &  2.13\\
+\hline
+2 & 3.03 &  0.13\\
+\hline
+3 & 10.19 & 2.28\\
+\hline
+4 & 9.88 &  0.03\\
+\hline
+\end{tabular}
+\caption{Number of targets hit and number of obstacles hit for human's behavior.}
+\end{table}
+
+In Figure~\ref{fig:heatmap}, we show the object values for different weights.
+
+\section{Conclusion and Future Work}
+
+Weighted sum of Q function is one way to combine multiple sub-MDPs. We also
+propose other ways, for example, scheduling between different modules, with only
+one active at one time. However, when a human tries to collect targets while
+avoiding obstacles, these two modules should be both on. A scheduling approach
+may yield frequent oscination between target and obstacle modules.
+
 \bibliographystyle{plain}

### 2015-02-11 paper.tex
exp results in

@@
 
+We conduct this research motivated by two questions.
+
 \begin{figure}[h!]
@@
 line on the ground, 2) getting targets, the blue cylinders, and 3) avoiding
-obstacles, the pink cylinders \cite{rothkopf2013modular}.
+obstacles, the pink cylinders. This is an experiment design used in the
+literature to evaluate modular reinforcement learning \cite{rothkopf2013modular}.
 
@@
 
+To obtain the weights given the samples, we need to use the Inverse Modular
+Reinforcement Learning technique \cite{rothkopf2013modular}.
+
 \section{Experiments}
@@
 simutanously.
+The human data are collected by the Center for Perceptual Systems in University
+of Texas at Austin.
 
@@
 red circles are obstacles. The blue circles are targets. The gray line is the
-path. The black lines are trajectories of human . The left figure shows the case
-where human follows the path and ignores the targets and obstacles. The right
-figure shows the case that human does three sub-tasks simultaneously. The human
-data are collected by the Center for Perceptual Systems in University of Texas
-at Austin.
+path. The black lines are trajectories of human.
+%TODO
 
-We make some constraints on our learning agent to make it behavior like humans.
-It is allowed to go straight ahead, turn left with a small step, or turn right
-with a small step.
+We make some constraints on our learning agent to make it walk like a human.
+We can find in the human trajectories that humans walk smoothly. They don't turn
+sharply.  Our agent is allowed to do three actions --- going straight ahead,
+turning left with a small step, and turning right with a small step.
 
-Now, we assume a learning agent that only knows the three sub-MDPs and the human
+To make our weights represent the significance of the modules, we
+normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
+is 1 for collecting a target, -1 for running into an obstacle. We define the
+value function directly for the path module to have a path following performance.
+
+The agent only considers the closest target and the closest obstacle.
+
+We assume that our learning agent only knows the three sub-MDPs and the human
 data. It looks at the human behavior, and finds the weights that can interpret
-such behavior. Using such weights, the trajectories of our agents are drawn in
-the green lines. We can tell that in the left figure, the agent puts a large
-weight on path-following. In the right figure, it puts weights on all sub-MDPs.
+such behavior. 
+Using such weights, the trajectories of our agents are drawn in the green lines.
+We can tell that in the left figure, the agent puts a large weight on
+path-following.  In the right figure, it puts weights on all sub-MDPs.
 
@@
 \hline
-2 & 3.03 &  0.13\\
+2 & 3.03 &  \bf{0.13}\\
 \hline
-3 & 10.19 & 2.28\\
+3 & \bf{10.19} & 2.28\\
 \hline
-4 & 9.88 &  0.03\\
+4 & \bf{9.88} &  \bf{0.03}\\
 \hline
 \end{tabular}
-\caption{Number of targets hit and number of obstacles hit for human's behavior.}
+\caption{Number of targets hit and number of obstacles hit of human.}
+\end{table}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{| l| l| l |}
+\hline
+Average by Task & Num Targs Hit & Num Obst Hit\\
+\hline
+1 & 1.25 & 1.62\\
+\hline
+2 & 3.62 & \bf{2.37}\\
+\hline
+3 & \bf{5.14} & 3.14\\
+\hline
+4 & \bf{5.00} & \bf{2.00}\\
+\hline
+\end{tabular}
+\caption{Number of targets hit and number of obstacles hit of the learning agent.}
 \end{table}
@@
 Weighted sum of Q function is one way to combine multiple sub-MDPs. We also
-propose other ways, for example, scheduling between different modules, with only
-one active at one time. However, when a human tries to collect targets while
-avoiding obstacles, these two modules should be both on. A scheduling approach
-may yield frequent oscination between target and obstacle modules.
+propose other ways including, for example, scheduling between different modules,
+with only one active at one time. However, we adopt the weighted sum approach
+because this is more reasonble for human behavior. When a human tries to collect
+targets while avoiding obstacles, these two modules are expected to be both
+active. A scheduling approach may yield frequent oscination between these two
+modules.
+
+We also assumes independency between modules. Correlation between modules
+doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap}, we can
+find that the target module and obstacle module tends to be negatively
+correlated.
 

### 2015-02-11 paper.tex
on rl

@@
 
-\title{Inverse Modular Reinforcement Learning on Human Motion}
+\title{Modular Inverse Reinforcement Learning on Human Motion}
 
@@
 \keywords{
-enter key words here
+Reinforcement learning, human cognition, inverse learning
 }
-
-\acknowledgements{}
 
@@
 
-Human is able to learn complicated behaviors much faster than machines can do.
+Human is able to learn to accomplish complicated tasks much faster than machines
+can do. However, most of the tasks can be decomposed into subtasks. A human may
+already have the capacity to accomplish the subtasks. He simply transfers the
+knowledge of the subtasks and integrate them for the objective task. For example,
+when a human learns to drive, he has much prior domain knowledge to help him
+with this objective task. He integrates his skills of controlling the velocity,
+avoiding other vehicles, navigating, responding to traffic lights, and so on.
 
-We analyze human's behavior of accomplishing various tasks. With the best of our
-knowledge, there is no work using inverse reinforcement learning to understand
-human motional behavior in the literature.
+Similar ideas are also adopted in the learning literature. Reinforcement learning
+suffers from the curse of dimensionality. It is inefficient for a learning agent
+to learn a complicated task from scratch. So learning algorithms including
+hierarchical reinforcement learning \cite{dietterich2000hierarchical}, modular
+reinforcement learning \cite{sprague2003multiple} are proposed. They also
+decompose the objective task into subtasks. The learning algorithms first learn
+the subtasks independently, then learn how to combine these subtasks.
 
-We conduct this research motivated by two questions.
+A natural question to ask is how human integrates skills for subtasks, and
+whether we can use the same method for integrating subtasks in reinforcement
+learning problems. In this paper, we analyze human's behavior of
+accomplishing a task composite of various subtasks. We collected human motion
+data when human accomplish a task, and try to interpret the behavior using
+inverse reinforcement learning. With the best of our knowledge, this approach is
+noval in the literature.
+
+This paper is organized as follows. Section~\ref{sec:domain} introduces the
+domain of the composite task that we collected human data. Section~\ref{sec:rl}
+describes the main algorithm, modular inverse reinforcement learning. We report
+our experiment results in Section~\ref{sec:exp}, and conclude in
+Section~\ref{sec:conclude}.
+
+\section{Multi-objective Sidewalk Domain}
+\label{sec:domain}
 
@@
 \caption{The task of collecting targets, avoiding obstacles and following the
-path.}
+path. \cite{rothkopf2013modular}}
 \label{fig:avatar}
@@
 
-Consider the task illustrated in Figure~\ref{fig:avatar} A). The avatar is asked to
-do three sub-tasks simultaneously --- 1) following the path, indicated by the gray
-line on the ground, 2) getting targets, the blue cylinders, and 3) avoiding
-obstacles, the pink cylinders. This is an experiment design used in the
-literature to evaluate modular reinforcement learning \cite{rothkopf2013modular}.
+Figure~\ref{fig:avatar} shows the domain that we use in this paper. The pink
+cylinders represent obstacles, that the player would be punished if running
+into. The blue cylinders represent targets, that the player would be rewarded if
+collected. There is also a grey path on the ground that the player can follow.
+So, naturally, this domain has three subtasks, 1) following the path, 2)
+collecting targets, 3) avoiding obstacles. This is an experiment used in the
+literature to evaluate modular reinforcement learning
+\cite{rothkopf2013modular}.
 
-From the reinforcement learning perspective, this task can be decomposed to be
-three sub-tasks as described above.  In Figure~\ref{fig:avatar} B), if the agent
-knows the distance and angle to an object, he is expected to know the optimal
-action to avoid or pursue it.
+We will evaluate four different tasks. Task 1, following the path only, and
+ignoring other objects. Task 2, following the path, while avoid the obstacles.
+Task 3, following the path, while attain targets when possible.  Task 4,
+following the path, collecting the targets and avoiding obstacles
+simultaneously.
+We conducted experiments to ask volunteers to accomplish different tasks and
+recorded their trajectories. The human data are collected by the Center for
+Perceptual Systems in University of Texas at Austin.
 
-A human has not done this kind of task before can achieve a good performance
-easily, shown in the experiments described later.
+If the player observes the distance and angle to an object, as shown in
+Figure~\ref{fig:avatar} B), he is expected to know the optimal action to avoid
+or pursue it. This decision is also Markov. Therefore, the question would be, if
+we have trained modules for these three subtasks, could they be integrated to
+match human's behavior?
 
-\section{Modular Reinforcement Learning}
+\section{Modular Inverse Reinforcement Learning}
+\label{sec:rl}
 
-Value functions is decomposed \cite{koller1999computing}.
+We need some basic concepts in reinforcement learning to proceed. For a state, action
+pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
+state $s$. The {\em policy} of a state is the action with the maximum Q
+value\cite{rl}. So how to we determine the global policy from the modules, or
+subtasks? In the literature, work has been done to integrate the Q function
+\cite{koller1999computing} or compromise on policies directly
+\cite{thomas2012motor}.
 
- If a human knows the policies
-of the sub-tasks, or sub-MDPs, he can accomplish a complicated behavior by
-combining the sub-MDPs. That is,
-$$Q(s, a) = \sum_i w_i Q_i (s, a)$$
-where $Q_i$ is the Q value of the i-th sub-MDP, $w_i$ is the weight of the i-th
-sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
+Here, we assume that the global Q function is a weighted sum of all $Q_i$, where
+$Q_i$ is the Q function for i-th module.
+$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
+where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
+$s_i$ denotes the decomposition of $s$ in the i-th module.
 
@@
 weights for the task of target collection, obstacle avoidance, and path
-following, respectively. Let $w$ be the vector of $(w_i)_1^n$. An agent with $w
-= (1, 0, 0)$ only collect targets, and one with $w = (0, 0.5, 0.5)$ may avoid
-the obstacles and follow the path.
+following, respectively. Let $w$ be the vector of $(w_1, w_2, w_3)$. An agent
+with $w = (1, 0, 0)$ only collect targets, and one with $w = (0, 0.5, 0.5)$ may
+avoid the obstacles and follow the path.
 
@@
 \section{Experiments}
-
-We conducted experiments to ask volunteers to accomplish different tasks and
-recorded their trajectories. There are four kind of tasks. Task 1, following the
-path only, and ignoring other objects. Task 2, following the path, while avoid
-the obstacles. Task 3, following the path, while attain targets when possible.
-Task 4, following the path, collecting the targets and avoiding obstacles
-simutanously.
-The human data are collected by the Center for Perceptual Systems in University
-of Texas at Austin.
+\label{sec:exp}
 
@@
 \end{tabular}
-\caption{Number of targets hit and number of obstacles hit of human.}
+\caption{Number of targets hit and number of obstacles hit of the humans.}
 \end{table}
@@
 \section{Conclusion and Future Work}
+\label{sec:conclude}
 
@@
 propose other ways including, for example, scheduling between different modules,
-with only one active at one time. However, we adopt the weighted sum approach
-because this is more reasonble for human behavior. When a human tries to collect
+with only one active at one time. This is also called skilled in the literature
+\cite{konidaris2009skill}. However, we adopt the weighted sum approach
+because this is more reasonable for human behavior. When a human tries to collect
 targets while avoiding obstacles, these two modules are expected to be both
-active. A scheduling approach may yield frequent oscination between these two
+active. A scheduling approach may yield frequent oscillation between these two
 modules.
@@
 
+Weights may be dynamic. However, with such assumption, we need to learn a
+mapping from state to weights. In this case, the curse of dimensionality still
+exists, and inverse learning would be difficult.
+
 \bibliographystyle{plain}

### 2015-02-11 paper.tex
experiment section

@@
 following, respectively. Let $w$ be the vector of $(w_1, w_2, w_3)$. An agent
-with $w = (1, 0, 0)$ only collect targets, and one with $w = (0, 0.5, 0.5)$ may
+with $w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
 avoid the obstacles and follow the path.
@@
 To obtain the weights given the samples, we need to use the Inverse Modular
-Reinforcement Learning technique \cite{rothkopf2013modular}.
+Reinforcement Learning technique \cite{rothkopf2013modular}. We use a maximum
+likelihood method here.
+\begin{equation}
+\label{eq:irl}
+\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
+\end{equation}
+where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
+$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
+before. $\eta$ is a hyperparameter that determines the consistency of human's
+behavior. The larger $\eta$ is, the algorithm more tends to overfit the data.
+
+The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
+sample, then the Q value of taking that action should be larger compared to Q
+values of taking other actions.
 
@@
 \label{sec:exp}
+
+We train the modules first to get $Q_i$ before running the inverse reinforcement
+learing algorithm. For each module, the agent only considers the closest target
+and the closest obstacle. For the path module, the path is defined as segments
+of waypoints, so the closest waypoint is considered. The agent takes the
+distance and angle to the closest objects as the state representation.
+
+To make our weights represent the significance of the modules, we
+normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
+is 1 for collecting a target, -1 for running into an obstacle. We define the
+value function directly for the path module to have a path following
+performance, as it is tricky to give reward for such performance.
+
+We make some constraints on our learning agent to make it walk like a human.
+We can find in the human trajectories that humans walk smoothly. They don't turn
+sharply.  Our agent is only allowed to do three actions --- going straight ahead,
+turning left with a small step, and turning right with a small step.
 
@@
 \includegraphics[width=0.24\textwidth]{task_4.png}
-\caption{}
+\caption{The trajectories of humans and our agent in four different tasks. The
+upper two are Task 1 and 2. The lower two are Task 3 and 4.}
 \label{fig:exp}
@@
 
-From a different perspective, can we find a weight vector to best interpret
-human's behavior. In Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar}, the
-red circles are obstacles. The blue circles are targets. The gray line is the
-path. The black lines are trajectories of human.
-%TODO
-
-We make some constraints on our learning agent to make it walk like a human.
-We can find in the human trajectories that humans walk smoothly. They don't turn
-sharply.  Our agent is allowed to do three actions --- going straight ahead,
-turning left with a small step, and turning right with a small step.
-
-To make our weights represent the significance of the modules, we
-normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
-is 1 for collecting a target, -1 for running into an obstacle. We define the
-value function directly for the path module to have a path following performance.
-
-The agent only considers the closest target and the closest obstacle.
-
-We assume that our learning agent only knows the three sub-MDPs and the human
-data. It looks at the human behavior, and finds the weights that can interpret
-such behavior. 
-Using such weights, the trajectories of our agents are drawn in the green lines.
-We can tell that in the left figure, the agent puts a large weight on
-path-following.  In the right figure, it puts weights on all sub-MDPs.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
-\caption{Heatmaps of the objective values of different weights for the four
-tasks, respectively. The red zones indicate higher values. The upper two are
-Task 1 and 2. The lower two are Task 3 and 4.}
-\label{fig:heatmap}
-\end{figure}
-
-\begin{table}[h]
+\begin{table}[h!]
 \centering
@@
 \caption{Number of targets hit and number of obstacles hit of the humans.}
+\label{tb:human}
 \end{table}
 
-\begin{table}[h]
+\begin{table}[h!]
 \centering
@@
 \caption{Number of targets hit and number of obstacles hit of the learning agent.}
+\label{tb:agent}
 \end{table}
 
-In Figure~\ref{fig:heatmap}, we show the object values for different weights.
+We report the results in Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar},
+the red circles are obstacles. The blue circles are targets. The gray line is
+the path. The black lines are trajectories of human --- each line represents one
+human trajectory.
+Using the weights derived from the algorithm, the trajectories of our agents are
+drawn in the green lines.
+
+Although humans don't agree each other on how to avoid obstacles or collecting
+targets, our agent can figure out what the humans are doing, and perform the
+similar performance. Table~\ref{tb:human} and Table~\ref{tb:agent} show the
+number of targets hit and number of obstacles hit. Note that the numbers bolded
+are active modules in the corresponding task. We can observe that humans still
+do better than our agent in these tasks.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
+\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
+\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
+\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
+\caption{Heatmaps of the $\log$ of values of Equation~\ref{eq:irl} for different
+weights for the four tasks, respectively. The red zones indicate higher values.
+From left to right are Task 1 to Task 4.}
+\label{fig:heatmap}
+\end{figure}
+
+In Figure~\ref{fig:heatmap}, we show the $\log$ of values of Equation~\ref{eq:irl} for
+different weights. We can observe the centroids of red zones move for different
+tasks. It stays at the origin in Task 1, so none of target and obstacle modules
+are active. It moves away from the origin when a module is active.
 
@@
 
-Weighted sum of Q function is one way to combine multiple sub-MDPs. We also
+Weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
 propose other ways including, for example, scheduling between different modules,

### 2015-02-11 paper.tex
changing while reading, sec3

@@
 
-Human is able to learn to accomplish complicated tasks much faster than machines
+Human is able to learn accomplishing complicated tasks much faster than machines
 can do. However, most of the tasks can be decomposed into subtasks. A human may
@@
 accomplishing a task composite of various subtasks. We collected human motion
-data when human accomplish a task, and try to interpret the behavior using
+data and try to interpret the behavior using a way of
 inverse reinforcement learning. With the best of our knowledge, this approach is
-noval in the literature.
+novel in the literature.
 
@@
 
-We will evaluate four different tasks. Task 1, following the path only, and
-ignoring other objects. Task 2, following the path, while avoid the obstacles.
-Task 3, following the path, while attain targets when possible.  Task 4,
+We will evaluate four different tasks. \bf{Task 1}, following the path only, and
+ignoring other objects. \bf{Task 2}, following the path, while avoid the obstacles.
+\bf{Task 3}, following the path, while attain targets when possible. \bf{Task 4},
 following the path, collecting the targets and avoiding obstacles
@@
 before. $\eta$ is a hyperparameter that determines the consistency of human's
-behavior. The larger $\eta$ is, the algorithm more tends to overfit the data.
+behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.
 
@@
 
+From the heatmaps, we find the optimums exist for these tasks. The optimal
+weights are also consistent with the experiment context.
+
 \section{Conclusion and Future Work}
@@
 
-Weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
+We interpreted human behavior using inverse module reinforcement learning in
+this paper. We have some positive results, but the performance of our agent is
+still inferior than humans.
+
+There are also some observations potential for the future work. First, 
+weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
 propose other ways including, for example, scheduling between different modules,
@@
 
-We also assumes independency between modules. Correlation between modules
-doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap}, we can
-find that the target module and obstacle module tends to be negatively
+Second, we also assumes independency between modules. Correlation between
+modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
+we can find that the target module and obstacle module tends to be negatively
 correlated.
 
-Weights may be dynamic. However, with such assumption, we need to learn a
-mapping from state to weights. In this case, the curse of dimensionality still
-exists, and inverse learning would be difficult.
+Lastly, weights may be dynamic and different from state to state. However, with
+such assumption, we need to learn a mapping from state to weights. In this case,
+the curse of dimensionality still exists, and inverse learning would be
+difficult.
 

### 2015-02-11 paper.tex
first version ok

@@
 \keywords{
-Reinforcement learning, human cognition, inverse learning
+Reinforcement learning, human motion, inverse learning
 }
@@
 into. The blue cylinders represent targets, that the player would be rewarded if
-collected. There is also a grey path on the ground that the player can follow.
+collected. There is also a gray path on the ground that the player can follow.
 So, naturally, this domain has three subtasks, 1) following the path, 2)
@@
 
-We will evaluate four different tasks. \bf{Task 1}, following the path only, and
-ignoring other objects. \bf{Task 2}, following the path, while avoid the obstacles.
-\bf{Task 3}, following the path, while attain targets when possible. \bf{Task 4},
+We will evaluate four different tasks. {\bf Task 1}, following the path only, and
+ignoring other objects. {\bf Task 2}, following the path, while avoid the obstacles.
+{\bf Task 3}, following the path, while attain targets when possible. {\bf Task 4},
 following the path, collecting the targets and avoiding obstacles
@@
 We train the modules first to get $Q_i$ before running the inverse reinforcement
-learing algorithm. For each module, the agent only considers the closest target
+learning algorithm. For each module, the agent only considers the closest target
 and the closest obstacle. For the path module, the path is defined as segments
@@
 \includegraphics[width=0.24\textwidth]{task_4.png}
-\caption{The trajectories of humans and our agent in four different tasks. The
-upper two are Task 1 and 2. The lower two are Task 3 and 4.}
+\caption{The trajectories of humans and our agent in four different tasks. From
+left to right are Task 1 to Task 4. The active modules should be (Path),
+(Obstacle + Path), (Target + Path), (Target + Obstacle + Path), respectively.}
 \label{fig:exp}
@@
 We report the results in Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar},
-the red circles are obstacles. The blue circles are targets. The gray line is
+the red circles are obstacles. The blue circles are targets. The gray lines are
 the path. The black lines are trajectories of human --- each line represents one
@@
 Using the weights derived from the algorithm, the trajectories of our agents are
-drawn in the green lines.
+drawn in green color.
 
 Although humans don't agree each other on how to avoid obstacles or collecting
-targets, our agent can figure out what the humans are doing, and perform the
-similar performance. Table~\ref{tb:human} and Table~\ref{tb:agent} show the
-number of targets hit and number of obstacles hit. Note that the numbers bolded
-are active modules in the corresponding task. We can observe that humans still
-do better than our agent in these tasks.
+targets, our agent can figure out what the humans are doing, and perform a
+similar trajectory. Table~\ref{tb:human} and Table~\ref{tb:agent} evaluate the
+performance by showing the number of targets hit and number of obstacles hit.
+Note that the numbers bold-ed are active modules in the corresponding task. We
+can observe that humans still do better than our agent in these tasks.
 
@@
 weights for the four tasks, respectively. The red zones indicate higher values.
-From left to right are Task 1 to Task 4.}
+The upper two are Task 1 and 2. The lower two are Task 3 and 4.
+}
 \label{fig:heatmap}
@@
 modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
-we can find that the target module and obstacle module tends to be negatively
-correlated.
+we can tell that the target module and obstacle module tend to be negatively
+correlated from the shape of the red zones.
 

### 2015-02-11 paper.tex
replaced avatar figure

@@
 \centering
-\includegraphics[width=0.6\textwidth]{avatar.png}
-\caption{The task of collecting targets, avoiding obstacles and following the
-path. \cite{rothkopf2013modular}}
+\includegraphics[height=6cm]{human.jpg}
+\includegraphics[height=6cm]{env.png}
+\caption{(Left) A volunteer with head tracker, virtual reality displayer, and
+body tracker.  (Right) The environment the human can see through the displayer.}
 \label{fig:avatar}
@@
 
-Figure~\ref{fig:avatar} shows the domain that we use in this paper. The pink
-cylinders represent obstacles, that the player would be punished if running
-into. The blue cylinders represent targets, that the player would be rewarded if
+Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
+cubes represent obstacles, that the player would be punished if running
+into. The blue balls represent targets, that the player would be rewarded if
 collected. There is also a gray path on the ground that the player can follow.
-So, naturally, this domain has three subtasks, 1) following the path, 2)
-collecting targets, 3) avoiding obstacles. This is an experiment used in the
+Naturally, this domain has three subtasks, 1) following the path, 2)
+collecting targets, 3) avoiding obstacles.
+This is an experiment used in the
 literature to evaluate modular reinforcement learning
 \cite{rothkopf2013modular}.
+We ask our volunteers to have trackers and virtual reality displayers equipped. He
+can see the envrionment through the displayer in front of his eyes. So he can
+walk as if he is walking in the virtual domain.
 
@@
 recorded their trajectories. The human data are collected by the Center for
-Perceptual Systems in University of Texas at Austin.
+Perceptual Systems at University of Texas at Austin.
 
-If the player observes the distance and angle to an object, as shown in
-Figure~\ref{fig:avatar} B), he is expected to know the optimal action to avoid
-or pursue it. This decision is also Markov. Therefore, the question would be, if
-we have trained modules for these three subtasks, could they be integrated to
-match human's behavior?
+If the player observes the distance and angle to an object, he is expected to
+know the optimal action to avoid or pursue it. This decision is also Markov.
+Therefore, the question would be, if we have trained modules for these three
+subtasks, could they be integrated to match human's behavior?
 
@@
 state $s$. The {\em policy} of a state is the action with the maximum Q
-value\cite{rl}. So how to we determine the global policy from the modules, or
+value \cite{rl}. So how to we determine the global policy from the modules, or
 subtasks? In the literature, work has been done to integrate the Q function

### 2015-02-11 paper.tex
set up for modification

@@
 
-Human is able to learn accomplishing complicated tasks much faster than machines
-can do. However, most of the tasks can be decomposed into subtasks. A human may
-already have the capacity to accomplish the subtasks. He simply transfers the
-knowledge of the subtasks and integrate them for the objective task. For example,
-when a human learns to drive, he has much prior domain knowledge to help him
-with this objective task. He integrates his skills of controlling the velocity,
-avoiding other vehicles, navigating, responding to traffic lights, and so on.
+A large body of evidence from human and other primate studies studies 
+suggests that they use reinforcement signals to learn tasks and reinforcement 
+learning has proved to be a good model for this process.
+However humans are able to learn and carry out very complex tasks involving 
+many different objectives.
 
-Similar ideas are also adopted in the learning literature. Reinforcement learning
-suffers from the curse of dimensionality. It is inefficient for a learning agent
-to learn a complicated task from scratch. So learning algorithms including
-hierarchical reinforcement learning \cite{dietterich2000hierarchical}, modular
-reinforcement learning \cite{sprague2003multiple} are proposed. They also
-decompose the objective task into subtasks. The learning algorithms first learn
-the subtasks independently, then learn how to combine these subtasks.
+How can they do this? The likelihood is that they use rinforcement in this 
+comtext as well but most formal reinforcement learning models do not scale 
+well with increasing complexity.
 
-A natural question to ask is how human integrates skills for subtasks, and
-whether we can use the same method for integrating subtasks in reinforcement
-learning problems. In this paper, we analyze human's behavior of
-accomplishing a task composite of various subtasks. We collected human motion
-data and try to interpret the behavior using a way of
-inverse reinforcement learning. With the best of our knowledge, this approach is
-novel in the literature.
+One promising possibiity is that the complex task can be broken down into 
+subtasks that are each learned separately \cite{sprague2003multiple,
+rothkopf2013modular}. 
+For certain task venues, this decomposition allows the behavior in the complex 
+task to be specified as a eighted sum of individal sub-task priorities.
+This paper explores the use of such a decomposition to characterize complex 
+behavior of subjects in a structure virtual reality environment. The 
+methodology of \cite{rothkopf2013modular} allows the task priorities to be
+estimated using the subjects' behavioral data.
+
+Our experimental analysis shows that the modular reinforcement can be a 
+surprisingly good model of behavior, predicting individual subjects' sub-task 
+priorities in a way that explains their beahvioral choices.
 
@@
 
-\section{Multi-objective Sidewalk Domain}
+\section{Multi-objective Domain}
 \label{sec:domain}
@@
 
-Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
+%FIXME
+In Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
 cubes represent obstacles, that the player would be punished if running
@@
 
-We need some basic concepts in reinforcement learning to proceed. For a state, action
+%FIXME
+For a state, action
 pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
@@
 
-We train the modules first to get $Q_i$ before running the inverse reinforcement
+We trained the modules first to get $Q_i$ before running the inverse reinforcement
 learning algorithm. For each module, the agent only considers the closest target
@@
 performance, as it is tricky to give reward for such performance.
+% emphasize that path following is not an RL module?
 

### 2015-02-11 paper.tex
fixed easy part first

@@
 \centering
-\includegraphics[height=6cm]{human.jpg}
-\includegraphics[height=6cm]{env.png}
-\caption{(Left) A volunteer with head tracker, virtual reality displayer, and
+\includegraphics[height=5cm]{human.jpg}
+\includegraphics[height=5cm]{env.png}
+\caption{(Left) A human subject with head tracker, virtual reality displayer, and
 body tracker.  (Right) The environment the human can see through the displayer.}
@@
 In Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
-cubes represent obstacles, that the player would be punished if running
-into. The blue balls represent targets, that the player would be rewarded if
-collected. There is also a gray path on the ground that the player can follow.
+cubes represent obstacles, that the agent would be punished if running
+into. The blue balls represent targets, that the agent would be rewarded if
+collected. There is also a gray path on the ground that the agent can follow.
 Naturally, this domain has three subtasks, 1) following the path, 2)
@@
 \cite{rothkopf2013modular}.
-We ask our volunteers to have trackers and virtual reality displayers equipped. He
+We ask human subjects to have trackers and virtual reality displayers equipped. He
 can see the envrionment through the displayer in front of his eyes. So he can
@@
 simultaneously.
-We conducted experiments to ask volunteers to accomplish different tasks and
+We conducted experiments to ask human subjects to accomplish different tasks and
 recorded their trajectories. The human data are collected by the Center for
@@
 
-If the player observes the distance and angle to an object, he is expected to
+If the agent observes the distance and angle to an object, he is expected to
 know the optimal action to avoid or pursue it. This decision is also Markov.
@@
 \includegraphics[width=0.24\textwidth]{task_4.png}
-\caption{The trajectories of humans and our agent in four different tasks. From
-left to right are Task 1 to Task 4. The active modules should be (Path),
+\caption{The trajectories of humans and the agent in four different tasks. (Path),
 (Obstacle + Path), (Target + Path), (Target + Obstacle + Path), respectively.}
@@
 \end{figure}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{| l| l| l |}
-\hline
-Average by Task & Num Targs Hit & Num Obst Hit\\
-\hline
-1 & 2.34 &  2.13\\
-\hline
-2 & 3.03 &  \bf{0.13}\\
-\hline
-3 & \bf{10.19} & 2.28\\
-\hline
-4 & \bf{9.88} &  \bf{0.03}\\
-\hline
-\end{tabular}
-\caption{Number of targets hit and number of obstacles hit of the humans.}
-\label{tb:human}
-\end{table}
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{| l| l| l |}
-\hline
-Average by Task & Num Targs Hit & Num Obst Hit\\
-\hline
-1 & 1.25 & 1.62\\
-\hline
-2 & 3.62 & \bf{2.37}\\
-\hline
-3 & \bf{5.14} & 3.14\\
-\hline
-4 & \bf{5.00} & \bf{2.00}\\
-\hline
-\end{tabular}
-\caption{Number of targets hit and number of obstacles hit of the learning agent.}
-\label{tb:agent}
-\end{table}
 
@@
 human trajectory.
-Using the weights derived from the algorithm, the trajectories of our agents are
+Using the weights derived from the algorithm, the trajectories of the agents are
 drawn in green color.
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.6\textwidth]{contactStats.png}
+\caption{Number of targets hit and number of obstacles hit of the human subjects
+and the agent.}
+\label{fig:heatmap}
+\end{figure}
+
 Although humans don't agree each other on how to avoid obstacles or collecting
-targets, our agent can figure out what the humans are doing, and perform a
+targets, the agent can figure out what the humans are doing, and perform a
 similar trajectory. Table~\ref{tb:human} and Table~\ref{tb:agent} evaluate the
@@
 Note that the numbers bold-ed are active modules in the corresponding task. We
-can observe that humans still do better than our agent in these tasks.
+can observe that humans still do better than the agent in these tasks.
 
@@
 We interpreted human behavior using inverse module reinforcement learning in
-this paper. We have some positive results, but the performance of our agent is
+this paper. We have some positive results, but the performance of the agent is
 still inferior than humans.

### 2015-02-12 paper.tex
subfigures

@@
 \usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 \caption{(Left) A human subject with head tracker, virtual reality displayer, and
-body tracker.  (Right) The environment the human can see through the displayer.}
+body tracker.  (Right) The environment the human can see through the displayer.
+The red cubes represent obstacles. The blue balls represent targets. There is
+also a gray path on the ground that the human subject can follow.}
 \label{fig:avatar}
@@
 
-%FIXME
-In Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
-cubes represent obstacles, that the agent would be punished if running
-into. The blue balls represent targets, that the agent would be rewarded if
-collected. There is also a gray path on the ground that the agent can follow.
+Figure~\ref{fig:avatar} shows the domain that we use for the experiment. 
+The agent would be punished if running into obstacles (red cubes), rewarded if
+collecting targets (blue balls). There is a path on the ground that he can
+follow.
 Naturally, this domain has three subtasks, 1) following the path, 2)
@@
 \centering
-\includegraphics[width=0.24\textwidth]{task_1.png}
-\includegraphics[width=0.24\textwidth]{task_2.png}
-\includegraphics[width=0.24\textwidth]{task_3.png}
-\includegraphics[width=0.24\textwidth]{task_4.png}
-\caption{The trajectories of humans and the agent in four different tasks. (Path),
-(Obstacle + Path), (Target + Path), (Target + Obstacle + Path), respectively.}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{task_1.png}
+\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
+\end{subfigure}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{task_2.png}
+\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
+\end{subfigure}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{task_3.png}
+\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
+\end{subfigure}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{task_4.png}
+\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
+\end{subfigure}
+\caption{The trajectories of humans and the agent in four different tasks. The
+black lines are trajectories of human subjects, and the green lines are
+trajectories of the learning agent by using the optimum weights, $w$. Weights
+are shown in target, obstacle, path order.}
+
 \label{fig:exp}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{contactStats.png}
+\includegraphics[width=0.5\textwidth]{contactStats.png}
 \caption{Number of targets hit and number of obstacles hit of the human subjects
 and the agent.}
-\label{fig:heatmap}
+\label{fig:stats}
 \end{figure}
@@
 targets, the agent can figure out what the humans are doing, and perform a
-similar trajectory. Table~\ref{tb:human} and Table~\ref{tb:agent} evaluate the
+similar trajectory. Figure~\ref{fig:stats} evaluates the
 performance by showing the number of targets hit and number of obstacles hit.
-Note that the numbers bold-ed are active modules in the corresponding task. We
-can observe that humans still do better than the agent in these tasks.
+We can observe that humans still do better than the agent in these tasks.
 
@@
 \centering
-\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
-\caption{Heatmaps of the $\log$ of values of Equation~\ref{eq:irl} for different
-weights for the four tasks, respectively. The red zones indicate higher values.
-The upper two are Task 1 and 2. The lower two are Task 3 and 4.
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{objValuesTask1.png}
+\caption{Path following only.}
+\end{subfigure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{objValuesTask2.png}
+\caption{Obstacle + Path. }
+\end{subfigure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{objValuesTask3.png}
+\caption{Target + Path. }
+\end{subfigure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{objValuesTask4.png}
+\caption{Target + Obstacle + Path. }
+\end{subfigure}
+\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for different
+weights for the four tasks, respectively. The white zones indicate higher
+values. The weights of all three modules sum to 1, so we only show the weights
+on the target and the obstacle modules.
 }
@@
 In Figure~\ref{fig:heatmap}, we show the $\log$ of values of Equation~\ref{eq:irl} for
-different weights. We can observe the centroids of red zones move for different
+different weights. We can observe the centroids of white zones move for different
 tasks. It stays at the origin in Task 1, so none of target and obstacle modules
@@
 we can tell that the target module and obstacle module tend to be negatively
-correlated from the shape of the red zones.
+correlated from the shape of the white zones.
 

### 2015-02-12 paper.tex
looks ok

@@
 
+%The abstract should be a maximum of 2000 characters of text, including
+%spaces (no figure is allowed).
 \begin{abstract}
-%The \emph{abstract} should be a maximum of 2000 characters of text, including
-%spaces (no figure is allowed).
+Human learning is modeled as reinforcement learning in the literature. However,
+human is able to learn complex tasks efficiently, while reinforcement
+learning is known for suffering from the curse of dimensionality. We observe that
+tasks that human handles can usually be decomposed into smaller tasks, and human can
+learn small tasks quickly, and then learn how to integrate them to solve the
+original task. In this paper, we find that modular reinforcement learning can be
+used to model human behavior. We report experiments of using modular inverse
+reinforcement learning to find the way that human combines sub-tasks, given
+samples of real human data. Our simulation demonstrates that human behavior can
+be well explained in this way.
 \end{abstract}
@@
 
-A large body of evidence from human and other primate studies studies 
-suggests that they use reinforcement signals to learn tasks and reinforcement 
+A large body of evidence from human and other primate studies 
+suggest that they use reinforcement signals to learn tasks and reinforcement 
 learning has proved to be a good model for this process.
@@
 many different objectives. 
-
-How can they do this? The likelihood is that they use rinforcement in this 
-comtext as well but most formal reinforcement learning models do not scale 
+How can they do this? The likelihood is that they use reinforcement in this 
+context as well but most formal reinforcement learning models do not scale 
 well with increasing complexity.
 
-One promising possibiity is that the complex task can be broken down into 
-subtasks that are each learned separately \cite{sprague2003multiple,
-rothkopf2013modular}. 
+One promising possibility is that the complex task can be broken down into 
+sub-tasks that are each learned separately \cite{sprague2003multiple,
+rothkopf2013modular, dietterich2000hierarchical}. 
 For certain task venues, this decomposition allows the behavior in the complex 
-task to be specified as a eighted sum of individal sub-task priorities.
+task to be specified as a weighted sum of individual sub-task priorities.
 This paper explores the use of such a decomposition to characterize complex
-behavior of subjects in a structure virtual reality environment. The 
-methodology of \cite{rothkopf2013modular} allows the task priorities to be
-estimated using the subjects' behavioral data.
+behavior of subjects in a virtual reality environment. The methodology of
+\cite{rothkopf2013modular} allows the task priorities to be estimated using the
+subjects' behavioral data.
 
@@
 surprisingly good model of behavior, predicting individual subjects' sub-task 
-priorities in a way that explains their beahvioral choices.
+priorities in a way that explains their behavioral choices.
 
@@
 
-Figure~\ref{fig:avatar} shows the domain that we use for the experiment. 
-The agent would be punished if running into obstacles (red cubes), rewarded if
-collecting targets (blue balls). There is a path on the ground that he can
+Figure~\ref{fig:avatar} shows the domain that we used for our experiment. The
+agent, which is a human subject in this case, is in a virtual reality domain.
+We ask the human subject to have trackers and a virtual reality displayer
+equipped. He can see the environment through the displayer in front of his eyes,
+and he can walk as if he is walking in the virtual domain.
+The agent would be punished if running into obstacles (red cubes), and rewarded
+if collecting targets (blue balls). There is a path on the ground that he can
 follow.
-Naturally, this domain has three subtasks, 1) following the path, 2)
+Naturally, this domain has three sub-tasks, 1) following the path, 2)
 collecting targets, 3) avoiding obstacles.
-This is an experiment used in the
-literature to evaluate modular reinforcement learning
-\cite{rothkopf2013modular}.
-We ask human subjects to have trackers and virtual reality displayers equipped. He
-can see the envrionment through the displayer in front of his eyes. So he can
-walk as if he is walking in the virtual domain.
+This is an experiment used in the literature to evaluate modular reinforcement
+learning \cite{rothkopf2013modular} and understand human behavior \cite{Tong2014}.
 
 We will evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoid the obstacles.
-{\bf Task 3}, following the path, while attain targets when possible. {\bf Task 4},
+ignoring other objects. {\bf Task 2}, following the path, while avoiding the obstacles.
+{\bf Task 3}, following the path, while collecting targets when possible. {\bf Task 4},
 following the path, collecting the targets and avoiding obstacles
 simultaneously.
-We conducted experiments to ask human subjects to accomplish different tasks and
+We conducted experiments to ask human subjects to accomplish these tasks and
 recorded their trajectories. The human data are collected by the Center for
@@
 Therefore, the question would be, if we have trained modules for these three
-subtasks, could they be integrated to match human's behavior?
+sub-tasks, could they be integrated to match human's behavior?
 
@@
 
-%FIXME
-For a state, action
+Recall that for a state, action
 pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
 state $s$. The {\em policy} of a state is the action with the maximum Q
-value \cite{rl}. So how to we determine the global policy from the modules, or
-subtasks? In the literature, work has been done to integrate the Q function
-\cite{koller1999computing} or compromise on policies directly
-\cite{thomas2012motor}.
+value \cite{rl}. So how do we determine the global policy from the modules, or
+sub-tasks? In the literature, work has been done to integrate the decomposed
+value functions of the sub-tasks \cite{koller1999computing}. The sub-tasks can
+also propose their own policies, while the global policy is a weighted sum of
+those sub-task policies \cite{thomas2012motor}.
 
@@
 
-Different weights can yield different performance. Let $w_1, w_2, w_3$ be
-weights for the task of target collection, obstacle avoidance, and path
-following, respectively. Let $w$ be the vector of $(w_1, w_2, w_3)$. An agent
-with $w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
+Different weights can yield different performance. Let $w$ be the vector of
+$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target
+collection, obstacle avoidance, and path following, respectively. An agent with
+$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
 avoid the obstacles and follow the path.
 
-To obtain the weights given the samples, we need to use the Inverse Modular
-Reinforcement Learning technique \cite{rothkopf2013modular}. We use a maximum
-likelihood method here.
+To obtain the weights given the samples, we need to use the inverse
+reinforcement learning technique. In reinforcement learning, we derive 
+optimal policies given an MDP. In inverse reinforcement learning, however, we
+find out the underlying MDP by observing policies. In the literature, a common
+way is to use a maximum likelihood method to recover the transition function and
+the reward function \cite{ng2000algorithms}. Here, however, we only need to find
+out the weights. The transition function is known, and the reward function is
+trivially the weighted sum of that of modules. Concretely, we use the modular
+inverse reinforcement learning method \cite{rothkopf2013modular} to maximize the
+function below.
 \begin{equation}
@@
 
-We trained the modules first to get $Q_i$ before running the inverse reinforcement
+We trained the modules first to obtain $Q_i$ before running the inverse reinforcement
 learning algorithm. For each module, the agent only considers the closest target
 and the closest obstacle. For the path module, the path is defined as segments
-of waypoints, so the closest waypoint is considered. The agent takes the
+connected by waypoints, so the next waypoint is considered. The agent takes the
 distance and angle to the closest objects as the state representation.
@@
 To make our weights represent the significance of the modules, we
-normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
-is 1 for collecting a target, -1 for running into an obstacle. We define the
-value function directly for the path module to have a path following
-performance, as it is tricky to give reward for such performance.
-% emphasize that path following is not an RL module?
+normalize the sub-MDPs with the unit (positive or negative) reward. The reward
+is 1 for collecting a target, -1 for running into an obstacle, and 1 for
+collecting a waypoint of the path. In the implementation, we make the radius of
+waypoints larger than that of other objects, so that the agent will not cling
+too much to the path.
 
 We make some constraints on our learning agent to make it walk like a human.
-We can find in the human trajectories that humans walk smoothly. They don't turn
+We can find from the human trajectories that humans walk smoothly and do not turn
 sharply.  Our agent is only allowed to do three actions --- going straight ahead,
@@
 
-\begin{figure}[h!]
+\begin{figure}[h]
 \centering
@@
 \end{subfigure}
-\caption{The trajectories of humans and the agent in four different tasks. The
+\caption{The trajectories of humans and the agent in the four tasks. The
 black lines are trajectories of human subjects, and the green lines are
-trajectories of the learning agent by using the optimum weights, $w$. Weights
-are shown in target, obstacle, path order.}
+trajectories of the learning agent by using the optimum weights, $w$, derived
+from modular inverse reinforcement learning. Weights are shown in the target,
+obstacle, path order.}
 
@@
 
-We report the results in Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar},
-the red circles are obstacles. The blue circles are targets. The gray lines are
-the path. The black lines are trajectories of human --- each line represents one
-human trajectory.
-Using the weights derived from the algorithm, the trajectories of the agents are
-drawn in green color.
-
-\begin{figure}[h!]
+\begin{figure}[h]
 \centering
@@
 
-Although humans don't agree each other on how to avoid obstacles or collecting
+We report the results in Figure~\ref{fig:exp}. Same as Figure~\ref{fig:avatar},
+the red circles are obstacles. The blue circles are targets. The gray lines are
+the path. The black lines are trajectories of human --- each line represents one
+human trajectory.
+Using the weights derived from the algorithm, the trajectories of the agents are
+drawn in green color.
+
+Although humans don't agree with each other on how to avoid obstacles or collect
 targets, the agent can figure out what the humans are doing, and perform a
-similar trajectory. Figure~\ref{fig:stats} evaluates the
+similar behavior. Figure~\ref{fig:stats} evaluates the
 performance by showing the number of targets hit and number of obstacles hit.
+The number of targets hit should be large, and the number of obstacles hit
+should be small, when the corresponding module is active.
 We can observe that humans still do better than the agent in these tasks.
+Nevertheless, the agent has a similar performance compared to human subjects.
 
-\begin{figure}[h!]
+\begin{figure}[h]
 \centering
@@
 \end{subfigure}
-\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for different
-weights for the four tasks, respectively. The white zones indicate higher
-values. The weights of all three modules sum to 1, so we only show the weights
-on the target and the obstacle modules.
+\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for
+different weights for the four tasks, respectively. The white zones indicate
+higher probabilities. The weights of all three modules sum to 1, so we only show
+the weights on the target and the obstacle modules.
 }
@@
 
-In Figure~\ref{fig:heatmap}, we show the $\log$ of values of Equation~\ref{eq:irl} for
-different weights. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so none of target and obstacle modules
-are active. It moves away from the origin when a module is active.
-
-From the heatmaps, we find the optimums exist for these tasks. The optimal
-weights are also consistent with the experiment context.
+In Figure~\ref{fig:heatmap}, we show the $\log$ of values of
+Equation~\ref{eq:irl} for different weights. The white color represents higher
+probability. We can observe the centroids of white zones move for different
+tasks. It stays at the origin in Task 1, so none of the target module and
+obstacle module is active. It moves away from the origin when a module is
+active.  From the heatmaps, we find the optimums exist for these tasks. The
+optimal weights are also consistent with the experiment context.
 
@@
 We interpreted human behavior using inverse module reinforcement learning in
-this paper. We have some positive results, but the performance of the agent is
-still inferior than humans.
+this paper. The experimental results show that modular reinforcement learning
+can well explain human behavior, even though the performance of the agent is
+inferior to human subjects'.
 
@@
 propose other ways including, for example, scheduling between different modules,
-with only one active at one time. This is also called skilled in the literature
+with only one active at one time. This is also called {\em skill} in the literature
 \cite{konidaris2009skill}. However, we adopt the weighted sum approach
@@
 
-Second, we also assumes independency between modules. Correlation between
+Second, we also assume independency between modules. Correlation between
 modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},

### 2015-02-14 content.tex
separated content from paper

--- 2015-02-12 paper.tex
@@
-\documentclass[11pt]{article} % For LaTeX2e
-\usepackage{rldmsubmit,palatino}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\title{Modular Inverse Reinforcement Learning on Human Motion}
-
-\author{
-Shun Zhang\\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{menie482@cs.utexas.edu} \\
-\And
-Matthew Tong \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{mhtong@gmail.com} \\
-\AND
-Mary Hayhoe \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{hayhoe@utexas.edu} \\
-\And
-Dana Ballard \\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{dana@cs.utexas.edu} \\
-\\
-}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to \LaTeX{} to determine where to break
-% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
-% puts 3 of 4 authors names on the first line, and the last on the second
-% line, try using \AND instead of \And before the third author name.
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\begin{document}
-
-\maketitle
-
-%The abstract should be a maximum of 2000 characters of text, including
-%spaces (no figure is allowed).
-\begin{abstract}
-Human learning is modeled as reinforcement learning in the literature. However,
-human is able to learn complex tasks efficiently, while reinforcement
-learning is known for suffering from the curse of dimensionality. We observe that
-tasks that human handles can usually be decomposed into smaller tasks, and human can
-learn small tasks quickly, and then learn how to integrate them to solve the
-original task. In this paper, we find that modular reinforcement learning can be
-used to model human behavior. We report experiments of using modular inverse
-reinforcement learning to find the way that human combines sub-tasks, given
-samples of real human data. Our simulation demonstrates that human behavior can
-be well explained in this way.
-\end{abstract}
-
-\keywords{
-Reinforcement learning, human motion, inverse learning
-}
-
-\startmain % to start the main 1-4 pages of the submission.
-
 \section{Introduction}
@@
 difficult.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2015-02-14 paper.tex
comment command
separated content from paper

--- 2015-02-11 paper.tex
@@
 
+\newcommand{\commentd}[1]{{\bf **Dana: #1**}}
+
 \begin{document}
@@
 
-\section{Introduction}
-
-A large body of evidence from human and other primate studies studies 
-suggests that they use reinforcement signals to learn tasks and reinforcement 
-learning has proved to be a good model for this process.
-However humans are able to learn and carry out very complex tasks involving 
-many different objectives.
-
-How can they do this? The likelihood is that they use rinforcement in this 
-comtext as well but most formal reinforcement learning models do not scale 
-well with increasing complexity.
-
-One promising possibiity is that the complex task can be broken down into 
-subtasks that are each learned separately \cite{sprague2003multiple,
-rothkopf2013modular}. 
-For certain task venues, this decomposition allows the behavior in the complex 
-task to be specified as a eighted sum of individal sub-task priorities.
-This paper explores the use of such a decomposition to characterize complex 
-behavior of subjects in a structure virtual reality environment. The 
-methodology of \cite{rothkopf2013modular} allows the task priorities to be
-estimated using the subjects' behavioral data.
-
-Our experimental analysis shows that the modular reinforcement can be a 
-surprisingly good model of behavior, predicting individual subjects' sub-task 
-priorities in a way that explains their beahvioral choices.
-
-This paper is organized as follows. Section~\ref{sec:domain} introduces the
-domain of the composite task that we collected human data. Section~\ref{sec:rl}
-describes the main algorithm, modular inverse reinforcement learning. We report
-our experiment results in Section~\ref{sec:exp}, and conclude in
-Section~\ref{sec:conclude}.
-
-\section{Multi-objective Domain}
-\label{sec:domain}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[height=5cm]{human.jpg}
-\includegraphics[height=5cm]{env.png}
-\caption{(Left) A human subject with head tracker, virtual reality displayer, and
-body tracker.  (Right) The environment the human can see through the displayer.}
-\label{fig:avatar}
-\end{figure}
-
-%FIXME
-In Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
-cubes represent obstacles, that the agent would be punished if running
-into. The blue balls represent targets, that the agent would be rewarded if
-collected. There is also a gray path on the ground that the agent can follow.
-Naturally, this domain has three subtasks, 1) following the path, 2)
-collecting targets, 3) avoiding obstacles.
-This is an experiment used in the
-literature to evaluate modular reinforcement learning
-\cite{rothkopf2013modular}.
-We ask human subjects to have trackers and virtual reality displayers equipped. He
-can see the envrionment through the displayer in front of his eyes. So he can
-walk as if he is walking in the virtual domain.
-
-We will evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoid the obstacles.
-{\bf Task 3}, following the path, while attain targets when possible. {\bf Task 4},
-following the path, collecting the targets and avoiding obstacles
-simultaneously.
-We conducted experiments to ask human subjects to accomplish different tasks and
-recorded their trajectories. The human data are collected by the Center for
-Perceptual Systems at University of Texas at Austin.
-
-If the agent observes the distance and angle to an object, he is expected to
-know the optimal action to avoid or pursue it. This decision is also Markov.
-Therefore, the question would be, if we have trained modules for these three
-subtasks, could they be integrated to match human's behavior?
-
-\section{Modular Inverse Reinforcement Learning}
-\label{sec:rl}
-
-%FIXME
-For a state, action
-pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
-state $s$. The {\em policy} of a state is the action with the maximum Q
-value \cite{rl}. So how to we determine the global policy from the modules, or
-subtasks? In the literature, work has been done to integrate the Q function
-\cite{koller1999computing} or compromise on policies directly
-\cite{thomas2012motor}.
-
-Here, we assume that the global Q function is a weighted sum of all $Q_i$, where
-$Q_i$ is the Q function for i-th module.
-$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
-where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-$s_i$ denotes the decomposition of $s$ in the i-th module.
-
-Different weights can yield different performance. Let $w_1, w_2, w_3$ be
-weights for the task of target collection, obstacle avoidance, and path
-following, respectively. Let $w$ be the vector of $(w_1, w_2, w_3)$. An agent
-with $w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
-avoid the obstacles and follow the path.
-
-To obtain the weights given the samples, we need to use the Inverse Modular
-Reinforcement Learning technique \cite{rothkopf2013modular}. We use a maximum
-likelihood method here.
-\begin{equation}
-\label{eq:irl}
-\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
-\end{equation}
-where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
-$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
-before. $\eta$ is a hyperparameter that determines the consistency of human's
-behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.
-
-The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
-sample, then the Q value of taking that action should be larger compared to Q
-values of taking other actions.
-
-\section{Experiments}
-\label{sec:exp}
-
-We trained the modules first to get $Q_i$ before running the inverse reinforcement
-learning algorithm. For each module, the agent only considers the closest target
-and the closest obstacle. For the path module, the path is defined as segments
-of waypoints, so the closest waypoint is considered. The agent takes the
-distance and angle to the closest objects as the state representation.
-
-To make our weights represent the significance of the modules, we
-normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
-is 1 for collecting a target, -1 for running into an obstacle. We define the
-value function directly for the path module to have a path following
-performance, as it is tricky to give reward for such performance.
-% emphasize that path following is not an RL module?
-
-We make some constraints on our learning agent to make it walk like a human.
-We can find in the human trajectories that humans walk smoothly. They don't turn
-sharply.  Our agent is only allowed to do three actions --- going straight ahead,
-turning left with a small step, and turning right with a small step.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.24\textwidth]{task_1.png}
-\includegraphics[width=0.24\textwidth]{task_2.png}
-\includegraphics[width=0.24\textwidth]{task_3.png}
-\includegraphics[width=0.24\textwidth]{task_4.png}
-\caption{The trajectories of humans and the agent in four different tasks. (Path),
-(Obstacle + Path), (Target + Path), (Target + Obstacle + Path), respectively.}
-\label{fig:exp}
-\end{figure}
-
-We report the results in Figure~\ref{fig:exp}, same as Figure~\ref{fig:avatar},
-the red circles are obstacles. The blue circles are targets. The gray lines are
-the path. The black lines are trajectories of human --- each line represents one
-human trajectory.
-Using the weights derived from the algorithm, the trajectories of the agents are
-drawn in green color.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.6\textwidth]{contactStats.png}
-\caption{Number of targets hit and number of obstacles hit of the human subjects
-and the agent.}
-\label{fig:heatmap}
-\end{figure}
-
-Although humans don't agree each other on how to avoid obstacles or collecting
-targets, the agent can figure out what the humans are doing, and perform a
-similar trajectory. Table~\ref{tb:human} and Table~\ref{tb:agent} evaluate the
-performance by showing the number of targets hit and number of obstacles hit.
-Note that the numbers bold-ed are active modules in the corresponding task. We
-can observe that humans still do better than the agent in these tasks.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.4\textwidth]{objValuesTask1.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask2.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask3.png}
-\includegraphics[width=0.4\textwidth]{objValuesTask4.png}
-\caption{Heatmaps of the $\log$ of values of Equation~\ref{eq:irl} for different
-weights for the four tasks, respectively. The red zones indicate higher values.
-The upper two are Task 1 and 2. The lower two are Task 3 and 4.
-}
-\label{fig:heatmap}
-\end{figure}
-
-In Figure~\ref{fig:heatmap}, we show the $\log$ of values of Equation~\ref{eq:irl} for
-different weights. We can observe the centroids of red zones move for different
-tasks. It stays at the origin in Task 1, so none of target and obstacle modules
-are active. It moves away from the origin when a module is active.
-
-From the heatmaps, we find the optimums exist for these tasks. The optimal
-weights are also consistent with the experiment context.
-
-\section{Conclusion and Future Work}
-\label{sec:conclude}
-
-We interpreted human behavior using inverse module reinforcement learning in
-this paper. We have some positive results, but the performance of the agent is
-still inferior than humans.
-
-There are also some observations potential for the future work. First, 
-weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
-propose other ways including, for example, scheduling between different modules,
-with only one active at one time. This is also called skilled in the literature
-\cite{konidaris2009skill}. However, we adopt the weighted sum approach
-because this is more reasonable for human behavior. When a human tries to collect
-targets while avoiding obstacles, these two modules are expected to be both
-active. A scheduling approach may yield frequent oscillation between these two
-modules.
-
-Second, we also assumes independency between modules. Correlation between
-modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
-we can tell that the target module and obstacle module tend to be negatively
-correlated from the shape of the red zones.
-
-Lastly, weights may be dynamic and different from state to state. However, with
-such assumption, we need to learn a mapping from state to weights. In this case,
-the curse of dimensionality still exists, and inverse learning would be
-difficult.
+\include{content}
 

### 2015-02-18 paper.tex
merged Mary's change

--- 2015-02-12 paper.tex
@@
 \begin{abstract}
-Human learning is modeled as reinforcement learning in the literature. However,
-human is able to learn complex tasks efficiently, while reinforcement
-learning is known for suffering from the curse of dimensionality. We observe that
-tasks that human handles can usually be decomposed into smaller tasks, and human can
-learn small tasks quickly, and then learn how to integrate them to solve the
-original task. In this paper, we find that modular reinforcement learning can be
-used to model human behavior. We report experiments of using modular inverse
-reinforcement learning to find the way that human combines sub-tasks, given
-samples of real human data. Our simulation demonstrates that human behavior can
-be well explained in this way.
+Reinforcement learning has been seen as a useful model for understanding human
+behavior because of the importance of the neural reward circuitry. However,
+because of the difficulty of scaling up RL models to large state spaces, it has
+been hard to apply these models to complex human behaviors. One potential
+simplification, consistent with observations of natural behavior, is that
+complex tasks can be broken down into independent sub-tasks, or modules . In
+this paper, we use observed human behavior while walking along a path to
+estimate the intrinsic reward values associated with different modular
+sub-tasks. To do this we use a simplified version of Inverse Reinforcement
+Learning to calculate the reward associated with path following, obstacle
+avoidance, and target collection of humans acting in an immersive virtual
+environment. Using the estimated values, a modular RL model can generate
+realistic behavior consistent with human action choices. This provides a way of
+understanding momentary sensorimotor decisions made in complex natural
+environments.
 \end{abstract}
@@
 For certain task venues, this decomposition allows the behavior in the complex 
-task to be specified as a weighted sum of individual sub-task priorities.
+task to be chosen based on the value of a weighted sum of individual sub-task priorities.
 This paper explores the use of such a decomposition to characterize complex
@@
 \cite{rothkopf2013modular} allows the task priorities to be estimated using the
-subjects' behavioral data.
-
-Our experimental analysis shows that the modular reinforcement can be a 
+subjects' behavioral data. Our experimental analysis shows that the modular reinforcement can be a 
 surprisingly good model of behavior, predicting individual subjects' sub-task 
@@
 \includegraphics[height=5cm]{env.png}
-\caption{(Left) A human subject with head tracker, virtual reality displayer, and
-body tracker.  (Right) The environment the human can see through the displayer.
+\caption{(Left) A human subject with head mounted display, eye, head, and
+body tracker.  (Right) The environment the human can see through the HMD.
 The red cubes represent obstacles. The blue balls represent targets. There is
@@
 agent, which is a human subject in this case, is in a virtual reality domain.
-We ask the human subject to have trackers and a virtual reality displayer
-equipped. He can see the environment through the displayer in front of his eyes,
-and he can walk as if he is walking in the virtual domain.
-The agent would be punished if running into obstacles (red cubes), and rewarded
-if collecting targets (blue balls). There is a path on the ground that he can
-follow.
-Naturally, this domain has three sub-tasks, 1) following the path, 2)
+The subject’s eye, head, and body motion was tracked as he/she walked through the virtual environment that was scaled to match the dimensions of a real environment.
+The subject was asked to follow the path, avoid the obstacles (red cubes), and collect the targets (blue balls) by intercepting them.
+In different conditions they were asked to give different importance to particular sub-tasks.
+Subjects were given auditory feedback when
+running into obstacles, and when intercepting targets, depending on their importance in the current condition. Thus this domain has three modules, 1) following the path, 2)
 collecting targets, 3) avoiding obstacles.
-This is an experiment used in the literature to evaluate modular reinforcement
+This paradigm has been used to evaluate modular reinforcement
 learning \cite{rothkopf2013modular} and understand human behavior \cite{Tong2014}.
@@
 simultaneously.
-We conducted experiments to ask human subjects to accomplish these tasks and
-recorded their trajectories. The human data are collected by the Center for
-Perceptual Systems at University of Texas at Austin.
-
+We conducted experiments using 4 human subjects who walked through the environment N times, in each of these conditions. 
 If the agent observes the distance and angle to an object, he is expected to
-know the optimal action to avoid or pursue it. This decision is also Markov.
-Therefore, the question would be, if we have trained modules for these three
-sub-tasks, could they be integrated to match human's behavior?
+know the optimal action to avoid or pursue it. We asked whether an agent with trained modules for these three
+sub-tasks could generate paths that matched human's behavior.
 
@@
 
-Recall that for a state, action
+In Reinforcement Learning with states and actions
 pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
@@
 learning algorithm. For each module, the agent only considers the closest target
-and the closest obstacle. For the path module, the path is defined as segments
-connected by waypoints, so the next waypoint is considered. The agent takes the
-distance and angle to the closest objects as the state representation.
-
-To make our weights represent the significance of the modules, we
-normalize the sub-MDPs with the unit (positive or negative) reward. The reward
+and the closest obstacle. For the path module, the path was defined as segments
+connected by waypoints, so only the next waypoint is considered. The agent takes the
+distance and angle to the closest objects as the state representation. To make our weights represent the significance of the modules, we
+normalized the sub-MDPs with the unit (positive or negative) reward. The reward
 is 1 for collecting a target, -1 for running into an obstacle, and 1 for
-collecting a waypoint of the path. In the implementation, we make the radius of
+collecting a waypoint of the path. In the implementation, we made the radius of
 waypoints larger than that of other objects, so that the agent will not cling
@@
 
-We make some constraints on our learning agent to make it walk like a human.
-We can find from the human trajectories that humans walk smoothly and do not turn
-sharply.  Our agent is only allowed to do three actions --- going straight ahead,
+We set some constraints on the learning agent in order to make it walk like a human.
+We observed from the human trajectories that humans walk smoothly and do not turn
+sharply.  Our agent was only allowed to do three actions --- going straight ahead,
 turning left with a small step, and turning right with a small step.
@@
 
-We report the results in Figure~\ref{fig:exp}. Same as Figure~\ref{fig:avatar},
-the red circles are obstacles. The blue circles are targets. The gray lines are
-the path. The black lines are trajectories of human --- each line represents one
+The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
+the red circles are obstacles and the blue circles are targets. The gray lines are
+the path. The black lines are trajectories of individual human subjects --- each line represents one
 human trajectory.
 Using the weights derived from the algorithm, the trajectories of the agents are
-drawn in green color.
-
-Although humans don't agree with each other on how to avoid obstacles or collect
-targets, the agent can figure out what the humans are doing, and perform a
-similar behavior. Figure~\ref{fig:stats} evaluates the
+drawn in green. Although human performance is variable on how to avoid obstacles or collect
+targets, the agent can figure out what the humans are doing, and perform 
+similarly. Figure~\ref{fig:stats} evaluates the
 performance by showing the number of targets hit and number of obstacles hit.
@@
 We can observe that humans still do better than the agent in these tasks.
-Nevertheless, the agent has a similar performance compared to human subjects.
+Nevertheless, the agent performance is quite similar to that of the human subjects.
 
@@
 obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find the optimums exist for these tasks. The
+active.  From the heatmaps, we find that fairly well-defined optima exist for these tasks. The
 optimal weights are also consistent with the experiment context.
 
-\section{Conclusion and Future Work}
+\section{Conclusions}
 \label{sec:conclude}
 
-We interpreted human behavior using inverse module reinforcement learning in
+We interpreted human behavior using inverse modular reinforcement learning in
 this paper. The experimental results show that modular reinforcement learning
@@
 
-There are also some observations potential for the future work. First, 
-weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
-propose other ways including, for example, scheduling between different modules,
+Note that a 
+weighted sum of Q functions is just one way to combine multiple sub-MDPs. Other ways are possible, including, for example, scheduling between different modules,
 with only one active at one time. This is also called {\em skill} in the literature
@@
 active. A scheduling approach may yield frequent oscillation between these two
-modules.
-
-Second, we also assume independency between modules. Correlation between
+modules. Note also that we assume independence between modules. Correlation between
 modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
 we can tell that the target module and obstacle module tend to be negatively
-correlated from the shape of the white zones.
-
-Lastly, weights may be dynamic and different from state to state. However, with
-such assumption, we need to learn a mapping from state to weights. In this case,
+correlated from the shape of the white zones. Lastly, weights may be dynamic and different from state to state. However, with
+such an assumption we need to learn a mapping from state to weights. In this case,
 the curse of dimensionality still exists, and inverse learning would be

### 2015-02-19 paper.tex
merged matt's change

@@
 How can they do this? The likelihood is that they use reinforcement in this 
-context as well but most formal reinforcement learning models do not scale 
+context as well, but most formal reinforcement learning models do not scale 
 well with increasing complexity.
@@
 This paper is organized as follows. Section~\ref{sec:domain} introduces the
-domain of the composite task that we collected human data. Section~\ref{sec:rl}
+domain and the composite task structure that we collected human data. Section~\ref{sec:rl}
 describes the main algorithm, modular inverse reinforcement learning. We report
@@
 \includegraphics[height=5cm]{env.png}
-\caption{(Left) A human subject with head mounted display, eye, head, and
-body tracker.  (Right) The environment the human can see through the HMD.
-The red cubes represent obstacles. The blue balls represent targets. There is
-also a gray path on the ground that the human subject can follow.}
+\caption{(Left) A human subject with a head mounted display (HMD) and trackers
+for the eye, head, and body.  (Right) The environment the human can see through
+the HMD.  The red cubes represent obstacles. The blue balls represent targets.
+There is also a gray path on the ground that the human subject can follow.}
 \label{fig:avatar}
@@
 Figure~\ref{fig:avatar} shows the domain that we used for our experiment. The
-agent, which is a human subject in this case, is in a virtual reality domain.
-The subject’s eye, head, and body motion was tracked as he/she walked through the virtual environment that was scaled to match the dimensions of a real environment.
-The subject was asked to follow the path, avoid the obstacles (red cubes), and collect the targets (blue balls) by intercepting them.
-In different conditions they were asked to give different importance to particular sub-tasks.
-Subjects were given auditory feedback when
-running into obstacles, and when intercepting targets, depending on their importance in the current condition. Thus this domain has three modules, 1) following the path, 2)
-collecting targets, 3) avoiding obstacles.
-This paradigm has been used to evaluate modular reinforcement
-learning \cite{rothkopf2013modular} and understand human behavior \cite{Tong2014}.
+agent, which was a human subject in this case, was in a virtual reality domain.
+The subject's eye, head, and body motion were tracked as he/she walked through a
+virtual environment that was designed to match the dimensions of the physical
+environment. The subject was asked to follow the path, avoid the obstacles (red
+cubes), and collect the targets (blue spheres) by intercepting them. In
+different conditions they were asked to give different importance to particular
+sub-tasks, with sub-tasks being either relevant or irrelevant. Subjects were
+given auditory feedback when running into obstacles, and when intercepting
+targets, depending on their importance in the current condition. Thus this
+domain has three modules, 1) following the path, 2) collecting targets, 3)
+avoiding obstacles.  This general paradigm has been used to evaluate modular
+reinforcement learning \cite{rothkopf2013modular} and understand human behavior
+\cite{Tong2014}.
 
-We will evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoiding the obstacles.
-{\bf Task 3}, following the path, while collecting targets when possible. {\bf Task 4},
-following the path, collecting the targets and avoiding obstacles
-simultaneously.
-We conducted experiments using 4 human subjects who walked through the environment N times, in each of these conditions. 
-If the agent observes the distance and angle to an object, he is expected to
-know the optimal action to avoid or pursue it. We asked whether an agent with trained modules for these three
-sub-tasks could generate paths that matched human's behavior.
+We evaluate four different tasks. {\bf Task 1}, following the path only, and
+ignoring other objects. {\bf Task 2}, following the path, while avoiding the
+obstacles.  {\bf Task 3}, following the path, while collecting targets. {\bf
+Task 4}, following the path, collecting the targets and avoiding obstacles
+simultaneously.  We conducted experiments using 4 human subjects who walked
+through the environment 8 times with different configurations of objects, in
+each of these 4 conditions.  We asked whether an agent with modules trained for
+these three sub-tasks could generate paths that matched human's behavior.
 
@@
 
-In Reinforcement Learning with states and actions
-pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
-state $s$. The {\em policy} of a state is the action with the maximum Q
-value \cite{rl}. So how do we determine the global policy from the modules, or
-sub-tasks? In the literature, work has been done to integrate the decomposed
-value functions of the sub-tasks \cite{koller1999computing}. The sub-tasks can
-also propose their own policies, while the global policy is a weighted sum of
-those sub-task policies \cite{thomas2012motor}.
+In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
+function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
+The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
+how do we determine the global policy from the modules, or sub-tasks? In the
+literature, work has been done to integrate the decomposed value functions of
+the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
+own policies, while the global policy is a weighted sum of those sub-task
+policies \cite{thomas2012motor}.
 
@@
 
-To obtain the weights given the samples, we need to use the inverse
-reinforcement learning technique. In reinforcement learning, we derive 
+To obtain the weights given the samples of observed human behavior, we can use inverse
+reinforcement learning. In reinforcement learning, we derive 
 optimal policies given an MDP. In inverse reinforcement learning, however, we
@@
 
-We trained the modules first to obtain $Q_i$ before running the inverse reinforcement
-learning algorithm. For each module, the agent only considers the closest target
-and the closest obstacle. For the path module, the path was defined as segments
-connected by waypoints, so only the next waypoint is considered. The agent takes the
-distance and angle to the closest objects as the state representation. To make our weights represent the significance of the modules, we
-normalized the sub-MDPs with the unit (positive or negative) reward. The reward
-is 1 for collecting a target, -1 for running into an obstacle, and 1 for
-collecting a waypoint of the path. In the implementation, we made the radius of
-waypoints larger than that of other objects, so that the agent will not cling
+We trained the modules first to obtain $Q_i$ before running the inverse
+reinforcement learning algorithm. For each module, the agent only considers the
+closest target and the closest obstacle. For the path module, the path was
+defined as segments connected by waypoints, so only the next waypoint is
+considered. The agent takes the distance and angle to the closest objects as the
+state representation. To make our weights represent the significance of the
+modules, we normalized the sub-MDPs with the unit (positive or negative) reward.
+The reward is 1 for collecting a target, -1 for running into an obstacle, and 1
+for collecting a waypoint of the path. In the implementation, we made the radius
+of waypoints larger than that of other objects, so that the agent will not cling
 too much to the path.
 
-We set some constraints on the learning agent in order to make it walk like a human.
-We observed from the human trajectories that humans walk smoothly and do not turn
-sharply.  Our agent was only allowed to do three actions --- going straight ahead,
-turning left with a small step, and turning right with a small step.
+We set some constraints on the learning agent in order to make it walk roughly
+like a human while keeping the action space small.  We observed from the human
+trajectories that humans walk smoothly and do not turn sharply.  Our agent was
+only allowed to do three actions --- going straight ahead, turning left with a
+small step, and turning right with a small step.
 
@@
 trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights are shown in the target,
-obstacle, path order.}
+from modular inverse reinforcement learning. Weights for each task are given as (target,
+obstacle, path).}
 
@@
 The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
-the red circles are obstacles and the blue circles are targets. The gray lines are
-the path. The black lines are trajectories of individual human subjects --- each line represents one
-human trajectory.
-Using the weights derived from the algorithm, the trajectories of the agents are
-drawn in green. Although human performance is variable on how to avoid obstacles or collect
-targets, the agent can figure out what the humans are doing, and perform 
-similarly. Figure~\ref{fig:stats} evaluates the
-performance by showing the number of targets hit and number of obstacles hit.
-The number of targets hit should be large, and the number of obstacles hit
-should be small, when the corresponding module is active.
-We can observe that humans still do better than the agent in these tasks.
-Nevertheless, the agent performance is quite similar to that of the human subjects.
+the red circles are obstacles and the blue circles are targets. The gray lines
+are the path. The black lines are trajectories of individual human subjects,
+with each line representing one human trajectory.  Using the weights recovered
+by IRL, the trajectories of the agents are drawn in green. Although individual
+subjects varied in how they avoided obstacles or collected targets, the agent
+can figure out what the humans are doing on aggregate, and perform similarly.
+Figure~\ref{fig:stats} evaluates the performance by showing the number of
+targets hit and number of obstacles hit.  The number of targets hit should be
+large, and the number of obstacles hit should be small, when the corresponding
+module is active.  We can observe that humans still do better than the agent in
+these tasks.  Nevertheless, the agent performance is quite similar to that of
+the human subjects.
 
@@
 probability. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so none of the target module and
+tasks. It stays at the origin in Task 1, so neither the target module nor the
 obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find that fairly well-defined optima exist for these tasks. The
-optimal weights are also consistent with the experiment context.
+active.  From the heatmaps, we find that fairly well-defined optima exist for
+these tasks. The optimal weights are also consistent with the experiment
+context.
 
@@
 
-We interpreted human behavior using inverse modular reinforcement learning in
-this paper. The experimental results show that modular reinforcement learning
-can well explain human behavior, even though the performance of the agent is
-inferior to human subjects'.
+We analyzed human behavior using inverse modular reinforcement learning. The
+experimental results show that modular reinforcement learning can explain human
+behavior well, even though the performance of the agent is currently inferior to
+human subjects'.
 
-Note that a 
-weighted sum of Q functions is just one way to combine multiple sub-MDPs. Other ways are possible, including, for example, scheduling between different modules,
-with only one active at one time. This is also called {\em skill} in the literature
-\cite{konidaris2009skill}. However, we adopt the weighted sum approach
-because this is more reasonable for human behavior. When a human tries to collect
-targets while avoiding obstacles, these two modules are expected to be both
-active. A scheduling approach may yield frequent oscillation between these two
-modules. Note also that we assume independence between modules. Correlation between
-modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
-we can tell that the target module and obstacle module tend to be negatively
-correlated from the shape of the white zones. Lastly, weights may be dynamic and different from state to state. However, with
-such an assumption we need to learn a mapping from state to weights. In this case,
-the curse of dimensionality still exists, and inverse learning would be
-difficult.
+Note that a weighted sum of Q functions is just one way to combine multiple
+sub-MDPs. Other ways are possible, including, for example, scheduling between
+different modules, with only one active at one time. This is also called {\em
+skill} in the literature \cite{konidaris2009skill}. However, we adopt the
+weighted sum approach because this is more reasonable for human behavior. When a
+human tries to collect targets while avoiding obstacles, these two modules are
+expected to be both active. A scheduling approach may yield frequent oscillation
+between these two modules. Note also that we assume independence between
+modules. However, correlation between modules doesn't impair our analysis in
+this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
+obstacle module tend to be negatively correlated from the shape of the white
+zones. Lastly, weights may be dynamic and different from state to state.
+However, with such an assumption we need to learn a mapping from state to
+weights. In this case, the curse of dimensionality still exists, and inverse
+learning would be difficult.
 

### 2015-02-20 old/paper.tex
respond to dana's

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{rldmsubmit,palatino}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{caption}
\usepackage{subcaption}

\title{Modular Inverse Reinforcement Learning on Human Motion}

\author{
Shun Zhang\\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{menie482@cs.utexas.edu} \\
\And
Matthew Tong \\
Center for Perceptual Systems\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{mhtong@gmail.com} \\
\AND
Mary Hayhoe \\
Center for Perceptual Systems\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{hayhoe@utexas.edu} \\
\And
Dana Ballard \\
Department of Computer Science\\
University of Texas at Austin\\
Austin, TX 78712 \\
\texttt{dana@cs.utexas.edu} \\
\\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\newcommand{\commentd}[1]{{\bf **Dana: #1**}}

\begin{document}

\maketitle

%The abstract should be a maximum of 2000 characters of text, including
%spaces (no figure is allowed).
\begin{abstract}
Reinforcement learning has been seen as a useful model for understanding human
behavior because of the importance of the neural reward circuitry. However,
because of the difficulty of scaling up RL models to large state spaces, it has
been hard to apply these models to complex human behaviors. One potential
simplification, consistent with observations of natural behavior, is that
complex tasks can be broken down into independent sub-tasks, or modules . In
this paper, we use observed human behavior while walking along a path to
estimate the intrinsic reward values associated with different modular
sub-tasks. To do this we use a simplified version of Inverse Reinforcement
Learning to calculate the reward associated with path following, obstacle
avoidance, and target collection of humans acting in an immersive virtual
environment. Using the estimated values, a modular RL model can generate
realistic behavior consistent with human action choices. This provides a way of
understanding momentary sensorimotor decisions made in complex natural
environments.
\end{abstract}

\keywords{
Reinforcement learning, human motion, inverse learning
}

\startmain % to start the main 1-4 pages of the submission.

\section{Introduction}

A large body of evidence from human and other primate studies 
suggests that they use reinforcement signals to learn tasks and reinforcement 
learning has proved to be a good model for this process.
However humans are able to learn and carry out very complex tasks involving 
many different objectives. 
How can they do this? The likelihood is that they use reinforcement in this 
context as well, but most formal reinforcement learning models do not scale 
well with increasing complexity.

One promising possibility is that the complex task can be broken down into 
sub-tasks that are each learned separately \cite{sprague2003multiple,
rothkopf2013modular, dietterich2000hierarchical}. 
For certain task venues, this decomposition allows the behavior in the complex 
task to be chosen based on the value of a weighted sum of individual sub-task priorities.
This paper explores the use of such a decomposition to characterize complex
behavior of subjects in a virtual reality environment. The methodology of
\cite{rothkopf2013modular} allows the task priorities to be estimated using the
subjects' behavioral data. Our experimental analysis shows that the modular reinforcement can be a 
surprisingly good model of behavior, predicting individual subjects' sub-task 
priorities in a way that explains their behavioral choices.

This paper is organized as follows. Section~\ref{sec:domain} introduces the
domain and the composite task structure that we collected human data. Section~\ref{sec:rl}
describes the main algorithm, modular inverse reinforcement learning. We report
our experiment results in Section~\ref{sec:exp}, and conclude in
Section~\ref{sec:conclude}.

\section{Multi-objective Domain}
\label{sec:domain}

\begin{SCfigure}[1.0][h]
\centering
\includegraphics[height=5cm]{human.jpg}
\includegraphics[height=6cm]{env.png}
\caption{(Left) A human subject with a head mounted display (HMD) and trackers
for the eye, head, and body.  (Right) The environment the human can see through
the HMD.  The red cubes represent obstacles. The blue balls represent targets.
There is also a gray path on the ground that the human subject can follow.}
\label{fig:avatar}
\end{SCfigure}

Figure~\ref{fig:avatar} shows the experimental domain. The
human subject was immersed in a virtual reality domain by wearing a binocular head-mounted display.
The subject's eye, head, and body motion were tracked as he/she walked through a
virtual environment that was designed to match the dimensions of a standard indoor
environment. The subject was asked to follow the path, avoid the obstacles (red
cubes), and collect the targets (blue spheres) by intercepting them. In
different conditions they were asked to give different importance to particular
sub-tasks, with sub-tasks being either relevant or irrelevant. Subjects were
given auditory feedback when running into obstacles, and when intercepting
targets, depending on their importance in the current condition. Thus this
domain can be modeled by three modules, 1) following the path, 2) collecting targets, 3)
avoiding obstacles.  This general paradigm has been used to evaluate modular
reinforcement learning \cite{rothkopf2013modular} and understand human behavior
\cite{Tong2014}.

We evaluate four different tasks. {\bf Task 1}, following the path only, and
ignoring other objects. {\bf Task 2}, following the path, while avoiding the
obstacles.  {\bf Task 3}, following the path, while collecting targets. {\bf
Task 4}, following the path, collecting the targets and avoiding obstacles
simultaneously.  We conducted experiments using 4 human subjects who walked
through the environment 8 times with different configurations of objects, in
each of these 4 conditions.  We asked whether an agent with modules trained for
these three sub-tasks could generate paths that matched human's behavior.

\section{Modular Inverse Reinforcement Learning}
\label{sec:rl}

In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
how do we determine the global policy from the modules, or sub-tasks?

%In the
%literature, work has been done to integrate the decomposed value functions of
%the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
%own policies, while the global policy is a weighted sum of those sub-task
%policies \cite{thomas2012motor}.
\textcolor{red}{COMMENT - DO NOT LEAVE IN Shun, the modular formulation started with Sprague et al ACM Trans. of Applied Perception 4(2), 1 , 2007 and this should be referenced.
The Parr and Koller paper is about factored MDPs which is diofferent than modular MDPs. The Thomas and Barto is a much more recent paper
that learns motor MDPs in a very restructed domain.}

Following (Sprague and Ballard 2007, Rothkopf and Ballard 2013), we assume that the global Q function is a weighted sum of all $Q_i$, where
$Q_i$ is the Q function for i-th module.
$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
$s_i$ denotes the decomposition of $s$ in the i-th module.

Different weights can yield different performance. Let $w$ be the vector of
$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target et al 
collection, obstacle avoidance, and path following, respectively. An agent with
$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
avoid the obstacles and follow the path.

To obtain the weights given the samples of observed human behavior, we use inverse
reinforcement learning. In reinforcement learning, we derive 
optimal policies given an MDP. In inverse reinforcement learning, however, aims to
find out the underlying MDP by observing policies. A common
way is to use a maximum likelihood method to recover the transition function and
the reward function \cite{ng2000algorithms}, but this approach is very expensive as some kind of learning
is used in the innermost search loop. The modular approach allows a much more economical approach (Rothkopf and Ballard), since we module weights
are the only free parameters. The transition function is known, and the reward function is
trivially the weighted sum of that of modules. More precisely, we use the modular
inverse reinforcement learning method in \cite{rothkopf2013modular} to maximize the
function below.
\begin{equation}
\label{eq:irl}
\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
\end{equation}
where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
before. $\eta$ is a hyperparameter that determines the consistency of human's
behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.

The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
sample, then the Q value of taking that action should be larger compared to Q
values of taking other actions.

\section{Experiments}
\label{sec:exp}

We trained the modules first to obtain $Q_i$ before running the inverse
reinforcement learning algorithm. For each module, the agent only considers the
closest target and the closest obstacle. For the path module, the path was
defined as segments connected by waypoints, so only the next waypoint is
considered. The agent takes the distance and angle to the closest objects as the
state representation. To make our weights represent the significance of the
modules, we normalized the sub-MDPs with the unit (positive or negative) reward.
The reward is 1 for collecting a target, -1 for running into an obstacle, and 1
for collecting a waypoint of the path. In the implementation, we made the radius
of waypoints larger than that of other objects, so that the agent will not cling
too much to the path.

We set some constraints on the learning agent in order to make it walk roughly
like a human while keeping the action space small.  We observed from the human
trajectories that humans walk smoothly and do not turn sharply.  Our agent was
only allowed to do three actions --- going straight ahead, turning left with a
small step, and turning right with a small step. \textcolor{red}{COMMENT It would be good to say how much in degrees}

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.24\textwidth}
\includegraphics[width=\textwidth]{task_1.png}
\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
\end{subfigure}
\begin{subfigure}[b]{0.24\textwidth}
\includegraphics[width=\textwidth]{task_2.png}
\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
\end{subfigure}
\begin{subfigure}[b]{0.24\textwidth}
\includegraphics[width=\textwidth]{task_3.png}
\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
\end{subfigure}
\begin{subfigure}[b]{0.24\textwidth}
\includegraphics[width=\textwidth]{task_4.png}
\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
\end{subfigure}
\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
black lines are trajectories of human subjects, and the green lines are
trajectories of the learning agent by using the optimum weights, $w$, derived
from modular inverse reinforcement learning. Weights for each task are given as (target,
obstacle, path).}

\label{fig:exp}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{contactStats.png}
\caption{Number of targets hit and number of obstacles hit of the human subjects 
and the agent.\textcolor{red}{COMMENT Shun it would be much better if you used black and green in this figure instead of blue and red since it agree 
with your previous figure. Also as a side comment its rather obvious that you could do better by using a path constraint that was not so severe}}
\label{fig:stats}
\end{figure}

The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
the red circles are obstacles and the blue circles are targets. The gray lines
are the path. The black lines are trajectories of individual human subjects,
with each line representing one human trajectory.  Using the weights recovered
by IRL, the trajectories of the agents are drawn in green. Although individual
subjects varied in how they avoided obstacles or collected targets, the agent
can figure out what the humans are doing on aggregate, and perform similarly.
Figure~\ref{fig:stats} evaluates the performance by showing the number of
targets hit and number of obstacles hit.  The number of targets hit should be
large, and the number of obstacles hit should be small, when the corresponding
module is active.  We can observe that humans still do better than the agent in
these tasks.  Nevertheless, the agent performance is quite similar to that of
the human subjects.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{objValuesTask1.png}
\caption{Path following only.}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{objValuesTask2.png}
\caption{Obstacle + Path. }
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{objValuesTask3.png}
\caption{Target + Path. }
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{objValuesTask4.png}
\caption{Target + Obstacle + Path. }
\end{subfigure}
\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for
different weights for the four tasks, respectively. The white zones indicate
higher probabilities. The weights of all three modules sum to 1, so we only show
the weights on the target and the obstacle modules.
}
\label{fig:heatmap}
\end{figure}

In Figure~\ref{fig:heatmap}, we show the $\log$ of values of
Equation~\ref{eq:irl} for different weights. The white color represents higher
probability. We can observe the centroids of white zones move for different
tasks. It stays at the origin in Task 1, so neither the target module nor the
obstacle module is active. It moves away from the origin when a module is
active.  From the heatmaps, we find that fairly well-defined optima exist for
these tasks. The optimal weights are also consistent with the experiment
context.

\section{Conclusions}
\label{sec:conclude}

We analyzed human behavior using inverse modular reinforcement learning. The
experimental results show that modular reinforcement learning can explain human
behavior well, even though the performance of the agent is currently inferior to
human subjects'.

Note that a weighted sum of Q functions is just one way to combine multiple
sub-MDPs. Other ways are possible, including, for example, scheduling between
different modules, with only one active at one time. This is also called {\em
skill} in the literature \cite{konidaris2009skill}. However, we adopt the
weighted sum approach because this is more reasonable for human behavior. When a
human tries to collect targets while avoiding obstacles, these two modules are
expected to be both active. A scheduling approach may yield frequent oscillation
between these two modules. Note also that we assume independence between
modules. However, correlation between modules doesn't impair our analysis in
this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
obstacle module tend to be negatively correlated from the shape of the white
zones. Lastly, weights may be dynamic and different from state to state.
However, with such an assumption we need to learn a mapping from state to
weights. In this case, the curse of dimensionality still exists, and inverse
learning would be difficult.

\bibliographystyle{plain}
\bibliography{paper}

\end{document}

### 2015-02-20 content.tex
use content.tex
updated content

@@
 A large body of evidence from human and other primate studies 
-suggest that they use reinforcement signals to learn tasks and reinforcement 
+suggests that they use reinforcement signals to learn tasks and reinforcement 
 learning has proved to be a good model for this process.
@@
 
-\begin{figure}[h!]
+\begin{figure}[h]
 \centering
 \includegraphics[height=5cm]{human.jpg}
-\includegraphics[height=5cm]{env.png}
-\caption{(Left) A human subject with head tracker, virtual reality displayer, and
-body tracker.  (Right) The environment the human can see through the displayer.
-The red cubes represent obstacles. The blue balls represent targets. There is
-also a gray path on the ground that the human subject can follow.}
+\includegraphics[height=6cm]{env.png}
+\caption{(Left) A human subject with a head mounted display (HMD) and trackers
+for the eye, head, and body.  (Right) The environment the human can see through
+the HMD.  The red cubes represent obstacles. The blue balls represent targets.
+There is also a gray path on the ground that the human subject can follow.}
 \label{fig:avatar}
@@
 
-Figure~\ref{fig:avatar} shows the domain that we used for our experiment. The
-agent, which is a human subject in this case, is in a virtual reality domain.
-We ask the human subject to have trackers and a virtual reality displayer
-equipped. He can see the environment through the displayer in front of his eyes,
-and he can walk as if he is walking in the virtual domain.
-The agent would be punished if running into obstacles (red cubes), and rewarded
-if collecting targets (blue balls). There is a path on the ground that he can
-follow.
-Naturally, this domain has three sub-tasks, 1) following the path, 2)
-collecting targets, 3) avoiding obstacles.
-This is an experiment used in the literature to evaluate modular reinforcement
-learning \cite{rothkopf2013modular} and understand human behavior \cite{Tong2014}.
+Figure~\ref{fig:avatar} shows the experimental domain. The
+human subject was immersed in a virtual reality domain by wearing a binocular head-mounted display.
+The subject's eye, head, and body motion were tracked as he/she walked through a
+virtual environment that was designed to match the dimensions of a standard indoor
+environment. The subject was asked to follow the path, avoid the obstacles (red
+cubes), and collect the targets (blue spheres) by intercepting them. In
+different conditions they were asked to give different importance to particular
+sub-tasks, with sub-tasks being either relevant or irrelevant. Subjects were
+given auditory feedback when running into obstacles, and when intercepting
+targets, depending on their importance in the current condition. Thus this
+domain can be modeled by three modules, 1) following the path, 2) collecting targets, 3)
+avoiding obstacles.  This general paradigm has been used to evaluate modular
+reinforcement learning \cite{rothkopf2013modular} and understand human behavior
+\cite{Tong2014}.
 
-We will evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoiding the obstacles.
-{\bf Task 3}, following the path, while collecting targets when possible. {\bf Task 4},
-following the path, collecting the targets and avoiding obstacles
-simultaneously.
-We conducted experiments to ask human subjects to accomplish these tasks and
-recorded their trajectories. The human data are collected by the Center for
-Perceptual Systems at University of Texas at Austin.
-
-If the agent observes the distance and angle to an object, he is expected to
-know the optimal action to avoid or pursue it. This decision is also Markov.
-Therefore, the question would be, if we have trained modules for these three
-sub-tasks, could they be integrated to match human's behavior?
+We evaluate four different tasks. {\bf Task 1}, following the path only, and
+ignoring other objects. {\bf Task 2}, following the path, while avoiding the
+obstacles.  {\bf Task 3}, following the path, while collecting targets. {\bf
+Task 4}, following the path, collecting the targets and avoiding obstacles
+simultaneously.  We conducted experiments using 4 human subjects who walked
+through the environment 8 times with different configurations of objects, in
+each of these 4 conditions.  We asked whether an agent with modules trained for
+these three sub-tasks could generate paths that matched human's behavior.
 
@@
 
-Recall that for a state, action
-pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
-state $s$. The {\em policy} of a state is the action with the maximum Q
-value \cite{rl}. So how do we determine the global policy from the modules, or
-sub-tasks? In the literature, work has been done to integrate the decomposed
-value functions of the sub-tasks \cite{koller1999computing}. The sub-tasks can
-also propose their own policies, while the global policy is a weighted sum of
-those sub-task policies \cite{thomas2012motor}.
+In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
+function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
+The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
+how do we determine the global policy from the modules, or sub-tasks?
 
-Here, we assume that the global Q function is a weighted sum of all $Q_i$, where
-$Q_i$ is the Q function for i-th module.
+%In the
+%literature, work has been done to integrate the decomposed value functions of
+%the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
+%own policies, while the global policy is a weighted sum of those sub-task
+%policies \cite{thomas2012motor}.
+
+%COMMENT: Shun, the modular formulation started with Sprague et al ACM Trans. of Applied Perception 4(2), 1 , 2007 and this should be referenced.
+%The Parr and Koller paper is about factored MDPs which is diofferent than modular MDPs. The Thomas and Barto is a much more recent paper
+%that learns motor MDPs in a very restructed domain.
+In the literature, modular formation is proposed for understanding human
+visuomotor behavior \cite{sprague2007modeling}. A combination of various basic
+modules can generate complex behavior to match human behavior.
+Following \cite{sprague2007modeling, rothkopf2013modular}, we consider how the
+modules can be combined. Here we assume that the global Q function is a weighted
+sum of all $Q_i$, where $Q_i$ is the Q function for i-th module.
 $$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
@@
 Different weights can yield different performance. Let $w$ be the vector of
-$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target
+$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target et al 
 collection, obstacle avoidance, and path following, respectively. An agent with
@@
 
-To obtain the weights given the samples, we need to use the inverse
-reinforcement learning technique. In reinforcement learning, we derive 
-optimal policies given an MDP. In inverse reinforcement learning, however, we
-find out the underlying MDP by observing policies. In the literature, a common
+To obtain the weights given the samples of observed human behavior, we use inverse
+reinforcement learning. In reinforcement learning, we derive 
+optimal policies given an MDP. In inverse reinforcement learning, however, aims to
+find out the underlying MDP by observing policies. A common
 way is to use a maximum likelihood method to recover the transition function and
-the reward function \cite{ng2000algorithms}. Here, however, we only need to find
-out the weights. The transition function is known, and the reward function is
-trivially the weighted sum of that of modules. Concretely, we use the modular
-inverse reinforcement learning method \cite{rothkopf2013modular} to maximize the
-function below.
+the reward function \cite{ng2000algorithms}, but this approach is very expensive
+as some kind of learning is used in the innermost search loop. The modular
+approach allows a much more economical approach \cite{rothkopf2013modular},
+since we module weights are the only free parameters. The transition function is
+known, and the reward function is trivially the weighted sum of that of modules.
+More precisely, we use the modular inverse reinforcement learning method in
+\cite{rothkopf2013modular} to maximize the function below.
 \begin{equation}
@@
 
-We trained the modules first to obtain $Q_i$ before running the inverse reinforcement
-learning algorithm. For each module, the agent only considers the closest target
-and the closest obstacle. For the path module, the path is defined as segments
-connected by waypoints, so the next waypoint is considered. The agent takes the
-distance and angle to the closest objects as the state representation.
-
-To make our weights represent the significance of the modules, we
-normalize the sub-MDPs with the unit (positive or negative) reward. The reward
-is 1 for collecting a target, -1 for running into an obstacle, and 1 for
-collecting a waypoint of the path. In the implementation, we make the radius of
-waypoints larger than that of other objects, so that the agent will not cling
+We trained the modules first to obtain $Q_i$ before running the inverse
+reinforcement learning algorithm. For each module, the agent only considers the
+closest target and the closest obstacle. For the path module, the path was
+defined as segments connected by waypoints, so only the next waypoint is
+considered. The agent takes the distance and angle to the closest objects as the
+state representation. To make our weights represent the significance of the
+modules, we normalized the sub-MDPs with the unit (positive or negative) reward.
+The reward is 1 for collecting a target, -1 for running into an obstacle, and 1
+for collecting a waypoint of the path. In the implementation, we made the radius
+of waypoints larger than that of other objects, so that the agent will not cling
 too much to the path.
 
-We make some constraints on our learning agent to make it walk like a human.
-We can find from the human trajectories that humans walk smoothly and do not turn
-sharply.  Our agent is only allowed to do three actions --- going straight ahead,
-turning left with a small step, and turning right with a small step.
+We set some constraints on the learning agent in order to make it walk roughly
+like a human while keeping the action space small.  We observed from the human
+trajectories that humans walk smoothly and do not turn sharply.  Our agent was
+only allowed to do three actions --- going straight ahead (0.3 meter),
+turning left with a small step (30 degrees to the left), and turning right with
+a small step (30 degrees to the right).
 
@@
 \end{subfigure}
-\caption{The trajectories of humans and the agent in the four tasks. The
+\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
 black lines are trajectories of human subjects, and the green lines are
 trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights are shown in the target,
-obstacle, path order.}
+from modular inverse reinforcement learning. Weights for each task are given as (target,
+obstacle, path).}
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{contactStats.png}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{contact1.png}
+\caption{Path module only. }
+\end{subfigure}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{contact2.png}
+\caption{Obstacle + Path. }
+\end{subfigure}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{contact3.png}
+\caption{Target + Path. }
+\end{subfigure}
+\begin{subfigure}[b]{0.24\textwidth}
+\includegraphics[width=\textwidth]{contact4.png}
+\caption{Target + Obstacle + Path. }
+\end{subfigure}
 \caption{Number of targets hit and number of obstacles hit of the human subjects 
 and the agent.}
+%COMMENT Shun it would be much better if you used black and green in this figure instead of blue and red since it agree 
+%with your previous figure. Also as a side comment its rather obvious that you could do better by using a path constraint that was not so severe
 \label{fig:stats}
@@
 
-We report the results in Figure~\ref{fig:exp}. Same as Figure~\ref{fig:avatar},
-the red circles are obstacles. The blue circles are targets. The gray lines are
-the path. The black lines are trajectories of human --- each line represents one
-human trajectory.
-Using the weights derived from the algorithm, the trajectories of the agents are
-drawn in green color.
-
-Although humans don't agree with each other on how to avoid obstacles or collect
-targets, the agent can figure out what the humans are doing, and perform a
-similar behavior. Figure~\ref{fig:stats} evaluates the
-performance by showing the number of targets hit and number of obstacles hit.
-The number of targets hit should be large, and the number of obstacles hit
-should be small, when the corresponding module is active.
-We can observe that humans still do better than the agent in these tasks.
-Nevertheless, the agent has a similar performance compared to human subjects.
+The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
+the red circles are obstacles and the blue circles are targets. The gray lines
+are the path. The black lines are trajectories of individual human subjects,
+with each line representing one human trajectory.  Using the weights recovered
+by IRL, the trajectories of the agents are drawn in green. Although individual
+subjects varied in how they avoided obstacles or collected targets, the agent
+can figure out what the humans are doing on aggregate, and perform similarly.
+Figure~\ref{fig:stats} evaluates the performance by showing the number of
+targets hit and number of obstacles hit.  The number of targets hit should be
+large, and the number of obstacles hit should be small, when the corresponding
+module is active.  We can observe that humans still do better than the agent in
+these tasks.  Nevertheless, the agent performance is quite similar to that of
+the human subjects.
 
@@
 probability. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so none of the target module and
+tasks. It stays at the origin in Task 1, so neither the target module nor the
 obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find the optimums exist for these tasks. The
-optimal weights are also consistent with the experiment context.
+active.  From the heatmaps, we find that fairly well-defined optima exist for
+these tasks. The optimal weights are also consistent with the experiment
+context.
 
-\section{Conclusion and Future Work}
+\section{Conclusions}
 \label{sec:conclude}
 
-We interpreted human behavior using inverse module reinforcement learning in
-this paper. The experimental results show that modular reinforcement learning
-can well explain human behavior, even though the performance of the agent is
-inferior to human subjects'.
+We analyzed human behavior using inverse modular reinforcement learning. The
+experimental results show that modular reinforcement learning can explain human
+behavior well, even though the performance of the agent is currently inferior to
+human subjects'.
 
-There are also some observations potential for the future work. First, 
-weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
-propose other ways including, for example, scheduling between different modules,
-with only one active at one time. This is also called {\em skill} in the literature
-\cite{konidaris2009skill}. However, we adopt the weighted sum approach
-because this is more reasonable for human behavior. When a human tries to collect
-targets while avoiding obstacles, these two modules are expected to be both
-active. A scheduling approach may yield frequent oscillation between these two
-modules.
-
-Second, we also assume independency between modules. Correlation between
-modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
-we can tell that the target module and obstacle module tend to be negatively
-correlated from the shape of the white zones.
-
-Lastly, weights may be dynamic and different from state to state. However, with
-such assumption, we need to learn a mapping from state to weights. In this case,
-the curse of dimensionality still exists, and inverse learning would be
-difficult.
+Note that a weighted sum of Q functions is just one way to combine multiple
+sub-MDPs. Other ways are possible, including, for example, scheduling between
+different modules, with only one active at one time. This is also called {\em
+skill} in the literature \cite{konidaris2009skill}. However, we adopt the
+weighted sum approach because this is more reasonable for human behavior. When a
+human tries to collect targets while avoiding obstacles, these two modules are
+expected to be both active. A scheduling approach may yield frequent oscillation
+between these two modules. Note also that we assume independence between
+modules. However, correlation between modules doesn't impair our analysis in
+this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
+obstacle module tend to be negatively correlated from the shape of the white
+zones. Lastly, weights may be dynamic and different from state to state.
+However, with such an assumption we need to learn a mapping from state to
+weights. In this case, the curse of dimensionality still exists, and inverse
+learning would be difficult.

### 2015-02-23 content.tex
update content

@@
 avoiding obstacles.  This general paradigm has been used to evaluate modular
-reinforcement learning \cite{rothkopf2013modular} and understand human behavior
+reinforcement learning \cite{Rothkopf12Infer, rothkopf2013modular} and understand human behavior
 \cite{Tong2014}.
@@
 visuomotor behavior \cite{sprague2007modeling}. A combination of various basic
-modules can generate complex behavior to match human behavior.
+modules can generate complex behavior to match human behavior.  The combination
+can be either of Q functions of modules \cite{rothkopf2013modular} or of rewards
+of modules \cite{Rothkopf12Infer}.
 Following \cite{sprague2007modeling, rothkopf2013modular}, we consider how the

### 2015-03-28 1-intro.tex
migrate to nips

--- 2015-02-20 old/paper.tex
@@
-\documentclass[11pt]{article} % For LaTeX2e
-\usepackage{rldmsubmit,palatino}
-\usepackage{graphicx}
-\usepackage{sidecap}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\title{Modular Inverse Reinforcement Learning on Human Motion}
-
-\author{
-Shun Zhang\\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{menie482@cs.utexas.edu} \\
-\And
-Matthew Tong \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{mhtong@gmail.com} \\
-\AND
-Mary Hayhoe \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{hayhoe@utexas.edu} \\
-\And
-Dana Ballard \\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{dana@cs.utexas.edu} \\
-\\
-}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to \LaTeX{} to determine where to break
-% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
-% puts 3 of 4 authors names on the first line, and the last on the second
-% line, try using \AND instead of \And before the third author name.
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\newcommand{\commentd}[1]{{\bf **Dana: #1**}}
-
-\begin{document}
-
-\maketitle
-
-%The abstract should be a maximum of 2000 characters of text, including
-%spaces (no figure is allowed).
-\begin{abstract}
-Reinforcement learning has been seen as a useful model for understanding human
-behavior because of the importance of the neural reward circuitry. However,
-because of the difficulty of scaling up RL models to large state spaces, it has
-been hard to apply these models to complex human behaviors. One potential
-simplification, consistent with observations of natural behavior, is that
-complex tasks can be broken down into independent sub-tasks, or modules . In
-this paper, we use observed human behavior while walking along a path to
-estimate the intrinsic reward values associated with different modular
-sub-tasks. To do this we use a simplified version of Inverse Reinforcement
-Learning to calculate the reward associated with path following, obstacle
-avoidance, and target collection of humans acting in an immersive virtual
-environment. Using the estimated values, a modular RL model can generate
-realistic behavior consistent with human action choices. This provides a way of
-understanding momentary sensorimotor decisions made in complex natural
-environments.
-\end{abstract}
-
-\keywords{
-Reinforcement learning, human motion, inverse learning
-}
-
-\startmain % to start the main 1-4 pages of the submission.
-
 \section{Introduction}
@@
 Section~\ref{sec:conclude}.
-
-\section{Multi-objective Domain}
-\label{sec:domain}
-
-\begin{SCfigure}[1.0][h]
-\centering
-\includegraphics[height=5cm]{human.jpg}
-\includegraphics[height=6cm]{env.png}
-\caption{(Left) A human subject with a head mounted display (HMD) and trackers
-for the eye, head, and body.  (Right) The environment the human can see through
-the HMD.  The red cubes represent obstacles. The blue balls represent targets.
-There is also a gray path on the ground that the human subject can follow.}
-\label{fig:avatar}
-\end{SCfigure}
-
-Figure~\ref{fig:avatar} shows the experimental domain. The
-human subject was immersed in a virtual reality domain by wearing a binocular head-mounted display.
-The subject's eye, head, and body motion were tracked as he/she walked through a
-virtual environment that was designed to match the dimensions of a standard indoor
-environment. The subject was asked to follow the path, avoid the obstacles (red
-cubes), and collect the targets (blue spheres) by intercepting them. In
-different conditions they were asked to give different importance to particular
-sub-tasks, with sub-tasks being either relevant or irrelevant. Subjects were
-given auditory feedback when running into obstacles, and when intercepting
-targets, depending on their importance in the current condition. Thus this
-domain can be modeled by three modules, 1) following the path, 2) collecting targets, 3)
-avoiding obstacles.  This general paradigm has been used to evaluate modular
-reinforcement learning \cite{rothkopf2013modular} and understand human behavior
-\cite{Tong2014}.
-
-We evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoiding the
-obstacles.  {\bf Task 3}, following the path, while collecting targets. {\bf
-Task 4}, following the path, collecting the targets and avoiding obstacles
-simultaneously.  We conducted experiments using 4 human subjects who walked
-through the environment 8 times with different configurations of objects, in
-each of these 4 conditions.  We asked whether an agent with modules trained for
-these three sub-tasks could generate paths that matched human's behavior.
-
-\section{Modular Inverse Reinforcement Learning}
-\label{sec:rl}
-
-In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
-function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
-The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
-how do we determine the global policy from the modules, or sub-tasks?
-
-%In the
-%literature, work has been done to integrate the decomposed value functions of
-%the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
-%own policies, while the global policy is a weighted sum of those sub-task
-%policies \cite{thomas2012motor}.
-\textcolor{red}{COMMENT - DO NOT LEAVE IN Shun, the modular formulation started with Sprague et al ACM Trans. of Applied Perception 4(2), 1 , 2007 and this should be referenced.
-The Parr and Koller paper is about factored MDPs which is diofferent than modular MDPs. The Thomas and Barto is a much more recent paper
-that learns motor MDPs in a very restructed domain.}
-
-Following (Sprague and Ballard 2007, Rothkopf and Ballard 2013), we assume that the global Q function is a weighted sum of all $Q_i$, where
-$Q_i$ is the Q function for i-th module.
-$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
-where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-$s_i$ denotes the decomposition of $s$ in the i-th module.
-
-Different weights can yield different performance. Let $w$ be the vector of
-$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target et al 
-collection, obstacle avoidance, and path following, respectively. An agent with
-$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
-avoid the obstacles and follow the path.
-
-To obtain the weights given the samples of observed human behavior, we use inverse
-reinforcement learning. In reinforcement learning, we derive 
-optimal policies given an MDP. In inverse reinforcement learning, however, aims to
-find out the underlying MDP by observing policies. A common
-way is to use a maximum likelihood method to recover the transition function and
-the reward function \cite{ng2000algorithms}, but this approach is very expensive as some kind of learning
-is used in the innermost search loop. The modular approach allows a much more economical approach (Rothkopf and Ballard), since we module weights
-are the only free parameters. The transition function is known, and the reward function is
-trivially the weighted sum of that of modules. More precisely, we use the modular
-inverse reinforcement learning method in \cite{rothkopf2013modular} to maximize the
-function below.
-\begin{equation}
-\label{eq:irl}
-\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
-\end{equation}
-where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
-$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
-before. $\eta$ is a hyperparameter that determines the consistency of human's
-behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.
-
-The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
-sample, then the Q value of taking that action should be larger compared to Q
-values of taking other actions.
-
-\section{Experiments}
-\label{sec:exp}
-
-We trained the modules first to obtain $Q_i$ before running the inverse
-reinforcement learning algorithm. For each module, the agent only considers the
-closest target and the closest obstacle. For the path module, the path was
-defined as segments connected by waypoints, so only the next waypoint is
-considered. The agent takes the distance and angle to the closest objects as the
-state representation. To make our weights represent the significance of the
-modules, we normalized the sub-MDPs with the unit (positive or negative) reward.
-The reward is 1 for collecting a target, -1 for running into an obstacle, and 1
-for collecting a waypoint of the path. In the implementation, we made the radius
-of waypoints larger than that of other objects, so that the agent will not cling
-too much to the path.
-
-We set some constraints on the learning agent in order to make it walk roughly
-like a human while keeping the action space small.  We observed from the human
-trajectories that humans walk smoothly and do not turn sharply.  Our agent was
-only allowed to do three actions --- going straight ahead, turning left with a
-small step, and turning right with a small step. \textcolor{red}{COMMENT It would be good to say how much in degrees}
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_1.png}
-\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_2.png}
-\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_3.png}
-\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_4.png}
-\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
-\end{subfigure}
-\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
-black lines are trajectories of human subjects, and the green lines are
-trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights for each task are given as (target,
-obstacle, path).}
-
-\label{fig:exp}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{contactStats.png}
-\caption{Number of targets hit and number of obstacles hit of the human subjects 
-and the agent.\textcolor{red}{COMMENT Shun it would be much better if you used black and green in this figure instead of blue and red since it agree 
-with your previous figure. Also as a side comment its rather obvious that you could do better by using a path constraint that was not so severe}}
-\label{fig:stats}
-\end{figure}
-
-The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
-the red circles are obstacles and the blue circles are targets. The gray lines
-are the path. The black lines are trajectories of individual human subjects,
-with each line representing one human trajectory.  Using the weights recovered
-by IRL, the trajectories of the agents are drawn in green. Although individual
-subjects varied in how they avoided obstacles or collected targets, the agent
-can figure out what the humans are doing on aggregate, and perform similarly.
-Figure~\ref{fig:stats} evaluates the performance by showing the number of
-targets hit and number of obstacles hit.  The number of targets hit should be
-large, and the number of obstacles hit should be small, when the corresponding
-module is active.  We can observe that humans still do better than the agent in
-these tasks.  Nevertheless, the agent performance is quite similar to that of
-the human subjects.
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask1.png}
-\caption{Path following only.}
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask2.png}
-\caption{Obstacle + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask3.png}
-\caption{Target + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask4.png}
-\caption{Target + Obstacle + Path. }
-\end{subfigure}
-\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for
-different weights for the four tasks, respectively. The white zones indicate
-higher probabilities. The weights of all three modules sum to 1, so we only show
-the weights on the target and the obstacle modules.
-}
-\label{fig:heatmap}
-\end{figure}
-
-In Figure~\ref{fig:heatmap}, we show the $\log$ of values of
-Equation~\ref{eq:irl} for different weights. The white color represents higher
-probability. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so neither the target module nor the
-obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find that fairly well-defined optima exist for
-these tasks. The optimal weights are also consistent with the experiment
-context.
-
-\section{Conclusions}
-\label{sec:conclude}
-
-We analyzed human behavior using inverse modular reinforcement learning. The
-experimental results show that modular reinforcement learning can explain human
-behavior well, even though the performance of the agent is currently inferior to
-human subjects'.
-
-Note that a weighted sum of Q functions is just one way to combine multiple
-sub-MDPs. Other ways are possible, including, for example, scheduling between
-different modules, with only one active at one time. This is also called {\em
-skill} in the literature \cite{konidaris2009skill}. However, we adopt the
-weighted sum approach because this is more reasonable for human behavior. When a
-human tries to collect targets while avoiding obstacles, these two modules are
-expected to be both active. A scheduling approach may yield frequent oscillation
-between these two modules. Note also that we assume independence between
-modules. However, correlation between modules doesn't impair our analysis in
-this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
-obstacle module tend to be negatively correlated from the shape of the white
-zones. Lastly, weights may be dynamic and different from state to state.
-However, with such an assumption we need to learn a mapping from state to
-weights. In this case, the curse of dimensionality still exists, and inverse
-learning would be difficult.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2015-03-28 2-domain.tex
migrate to nips

--- 2015-02-23 content.tex
@@
-\section{Introduction}
-
-A large body of evidence from human and other primate studies 
-suggests that they use reinforcement signals to learn tasks and reinforcement 
-learning has proved to be a good model for this process.
-However humans are able to learn and carry out very complex tasks involving 
-many different objectives. 
-How can they do this? The likelihood is that they use reinforcement in this 
-context as well, but most formal reinforcement learning models do not scale 
-well with increasing complexity.
-
-One promising possibility is that the complex task can be broken down into 
-sub-tasks that are each learned separately \cite{sprague2003multiple,
-rothkopf2013modular, dietterich2000hierarchical}. 
-For certain task venues, this decomposition allows the behavior in the complex 
-task to be chosen based on the value of a weighted sum of individual sub-task priorities.
-This paper explores the use of such a decomposition to characterize complex
-behavior of subjects in a virtual reality environment. The methodology of
-\cite{rothkopf2013modular} allows the task priorities to be estimated using the
-subjects' behavioral data. Our experimental analysis shows that the modular reinforcement can be a 
-surprisingly good model of behavior, predicting individual subjects' sub-task 
-priorities in a way that explains their behavioral choices.
-
-This paper is organized as follows. Section~\ref{sec:domain} introduces the
-domain and the composite task structure that we collected human data. Section~\ref{sec:rl}
-describes the main algorithm, modular inverse reinforcement learning. We report
-our experiment results in Section~\ref{sec:exp}, and conclude in
-Section~\ref{sec:conclude}.
-
 \section{Multi-objective Domain}
@@
 these three sub-tasks could generate paths that matched human's behavior.
-
-\section{Modular Inverse Reinforcement Learning}
-\label{sec:rl}
-
-In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
-function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
-The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
-how do we determine the global policy from the modules, or sub-tasks?
-
-%In the
-%literature, work has been done to integrate the decomposed value functions of
-%the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
-%own policies, while the global policy is a weighted sum of those sub-task
-%policies \cite{thomas2012motor}.
-
-%COMMENT: Shun, the modular formulation started with Sprague et al ACM Trans. of Applied Perception 4(2), 1 , 2007 and this should be referenced.
-%The Parr and Koller paper is about factored MDPs which is diofferent than modular MDPs. The Thomas and Barto is a much more recent paper
-%that learns motor MDPs in a very restructed domain.
-In the literature, modular formation is proposed for understanding human
-visuomotor behavior \cite{sprague2007modeling}. A combination of various basic
-modules can generate complex behavior to match human behavior.  The combination
-can be either of Q functions of modules \cite{rothkopf2013modular} or of rewards
-of modules \cite{Rothkopf12Infer}.
-Following \cite{sprague2007modeling, rothkopf2013modular}, we consider how the
-modules can be combined. Here we assume that the global Q function is a weighted
-sum of all $Q_i$, where $Q_i$ is the Q function for i-th module.
-$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
-where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-$s_i$ denotes the decomposition of $s$ in the i-th module.
-
-Different weights can yield different performance. Let $w$ be the vector of
-$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target et al 
-collection, obstacle avoidance, and path following, respectively. An agent with
-$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
-avoid the obstacles and follow the path.
-
-To obtain the weights given the samples of observed human behavior, we use inverse
-reinforcement learning. In reinforcement learning, we derive 
-optimal policies given an MDP. In inverse reinforcement learning, however, aims to
-find out the underlying MDP by observing policies. A common
-way is to use a maximum likelihood method to recover the transition function and
-the reward function \cite{ng2000algorithms}, but this approach is very expensive
-as some kind of learning is used in the innermost search loop. The modular
-approach allows a much more economical approach \cite{rothkopf2013modular},
-since we module weights are the only free parameters. The transition function is
-known, and the reward function is trivially the weighted sum of that of modules.
-More precisely, we use the modular inverse reinforcement learning method in
-\cite{rothkopf2013modular} to maximize the function below.
-\begin{equation}
-\label{eq:irl}
-\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
-\end{equation}
-where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
-$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
-before. $\eta$ is a hyperparameter that determines the consistency of human's
-behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.
-
-The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
-sample, then the Q value of taking that action should be larger compared to Q
-values of taking other actions.
-
-\section{Experiments}
-\label{sec:exp}
-
-We trained the modules first to obtain $Q_i$ before running the inverse
-reinforcement learning algorithm. For each module, the agent only considers the
-closest target and the closest obstacle. For the path module, the path was
-defined as segments connected by waypoints, so only the next waypoint is
-considered. The agent takes the distance and angle to the closest objects as the
-state representation. To make our weights represent the significance of the
-modules, we normalized the sub-MDPs with the unit (positive or negative) reward.
-The reward is 1 for collecting a target, -1 for running into an obstacle, and 1
-for collecting a waypoint of the path. In the implementation, we made the radius
-of waypoints larger than that of other objects, so that the agent will not cling
-too much to the path.
-
-We set some constraints on the learning agent in order to make it walk roughly
-like a human while keeping the action space small.  We observed from the human
-trajectories that humans walk smoothly and do not turn sharply.  Our agent was
-only allowed to do three actions --- going straight ahead (0.3 meter),
-turning left with a small step (30 degrees to the left), and turning right with
-a small step (30 degrees to the right).
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_1.png}
-\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_2.png}
-\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_3.png}
-\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_4.png}
-\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
-\end{subfigure}
-\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
-black lines are trajectories of human subjects, and the green lines are
-trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights for each task are given as (target,
-obstacle, path).}
-
-\label{fig:exp}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact1.png}
-\caption{Path module only. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact2.png}
-\caption{Obstacle + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact3.png}
-\caption{Target + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact4.png}
-\caption{Target + Obstacle + Path. }
-\end{subfigure}
-\caption{Number of targets hit and number of obstacles hit of the human subjects 
-and the agent.}
-%COMMENT Shun it would be much better if you used black and green in this figure instead of blue and red since it agree 
-%with your previous figure. Also as a side comment its rather obvious that you could do better by using a path constraint that was not so severe
-\label{fig:stats}
-\end{figure}
-
-The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
-the red circles are obstacles and the blue circles are targets. The gray lines
-are the path. The black lines are trajectories of individual human subjects,
-with each line representing one human trajectory.  Using the weights recovered
-by IRL, the trajectories of the agents are drawn in green. Although individual
-subjects varied in how they avoided obstacles or collected targets, the agent
-can figure out what the humans are doing on aggregate, and perform similarly.
-Figure~\ref{fig:stats} evaluates the performance by showing the number of
-targets hit and number of obstacles hit.  The number of targets hit should be
-large, and the number of obstacles hit should be small, when the corresponding
-module is active.  We can observe that humans still do better than the agent in
-these tasks.  Nevertheless, the agent performance is quite similar to that of
-the human subjects.
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask1.png}
-\caption{Path following only.}
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask2.png}
-\caption{Obstacle + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask3.png}
-\caption{Target + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask4.png}
-\caption{Target + Obstacle + Path. }
-\end{subfigure}
-\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for
-different weights for the four tasks, respectively. The white zones indicate
-higher probabilities. The weights of all three modules sum to 1, so we only show
-the weights on the target and the obstacle modules.
-}
-\label{fig:heatmap}
-\end{figure}
-
-In Figure~\ref{fig:heatmap}, we show the $\log$ of values of
-Equation~\ref{eq:irl} for different weights. The white color represents higher
-probability. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so neither the target module nor the
-obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find that fairly well-defined optima exist for
-these tasks. The optimal weights are also consistent with the experiment
-context.
-
-\section{Conclusions}
-\label{sec:conclude}
-
-We analyzed human behavior using inverse modular reinforcement learning. The
-experimental results show that modular reinforcement learning can explain human
-behavior well, even though the performance of the agent is currently inferior to
-human subjects'.
-
-Note that a weighted sum of Q functions is just one way to combine multiple
-sub-MDPs. Other ways are possible, including, for example, scheduling between
-different modules, with only one active at one time. This is also called {\em
-skill} in the literature \cite{konidaris2009skill}. However, we adopt the
-weighted sum approach because this is more reasonable for human behavior. When a
-human tries to collect targets while avoiding obstacles, these two modules are
-expected to be both active. A scheduling approach may yield frequent oscillation
-between these two modules. Note also that we assume independence between
-modules. However, correlation between modules doesn't impair our analysis in
-this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
-obstacle module tend to be negatively correlated from the shape of the white
-zones. Lastly, weights may be dynamic and different from state to state.
-However, with such an assumption we need to learn a mapping from state to
-weights. In this case, the curse of dimensionality still exists, and inverse
-learning would be difficult.

### 2015-03-28 3-irl.tex
migrate to nips

--- 2015-02-23 content.tex
@@
-\section{Introduction}
-
-A large body of evidence from human and other primate studies 
-suggests that they use reinforcement signals to learn tasks and reinforcement 
-learning has proved to be a good model for this process.
-However humans are able to learn and carry out very complex tasks involving 
-many different objectives. 
-How can they do this? The likelihood is that they use reinforcement in this 
-context as well, but most formal reinforcement learning models do not scale 
-well with increasing complexity.
-
-One promising possibility is that the complex task can be broken down into 
-sub-tasks that are each learned separately \cite{sprague2003multiple,
-rothkopf2013modular, dietterich2000hierarchical}. 
-For certain task venues, this decomposition allows the behavior in the complex 
-task to be chosen based on the value of a weighted sum of individual sub-task priorities.
-This paper explores the use of such a decomposition to characterize complex
-behavior of subjects in a virtual reality environment. The methodology of
-\cite{rothkopf2013modular} allows the task priorities to be estimated using the
-subjects' behavioral data. Our experimental analysis shows that the modular reinforcement can be a 
-surprisingly good model of behavior, predicting individual subjects' sub-task 
-priorities in a way that explains their behavioral choices.
-
-This paper is organized as follows. Section~\ref{sec:domain} introduces the
-domain and the composite task structure that we collected human data. Section~\ref{sec:rl}
-describes the main algorithm, modular inverse reinforcement learning. We report
-our experiment results in Section~\ref{sec:exp}, and conclude in
-Section~\ref{sec:conclude}.
-
-\section{Multi-objective Domain}
-\label{sec:domain}
-
-\begin{figure}[h]
-\centering
-\includegraphics[height=5cm]{human.jpg}
-\includegraphics[height=6cm]{env.png}
-\caption{(Left) A human subject with a head mounted display (HMD) and trackers
-for the eye, head, and body.  (Right) The environment the human can see through
-the HMD.  The red cubes represent obstacles. The blue balls represent targets.
-There is also a gray path on the ground that the human subject can follow.}
-\label{fig:avatar}
-\end{figure}
-
-Figure~\ref{fig:avatar} shows the experimental domain. The
-human subject was immersed in a virtual reality domain by wearing a binocular head-mounted display.
-The subject's eye, head, and body motion were tracked as he/she walked through a
-virtual environment that was designed to match the dimensions of a standard indoor
-environment. The subject was asked to follow the path, avoid the obstacles (red
-cubes), and collect the targets (blue spheres) by intercepting them. In
-different conditions they were asked to give different importance to particular
-sub-tasks, with sub-tasks being either relevant or irrelevant. Subjects were
-given auditory feedback when running into obstacles, and when intercepting
-targets, depending on their importance in the current condition. Thus this
-domain can be modeled by three modules, 1) following the path, 2) collecting targets, 3)
-avoiding obstacles.  This general paradigm has been used to evaluate modular
-reinforcement learning \cite{Rothkopf12Infer, rothkopf2013modular} and understand human behavior
-\cite{Tong2014}.
-
-We evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoiding the
-obstacles.  {\bf Task 3}, following the path, while collecting targets. {\bf
-Task 4}, following the path, collecting the targets and avoiding obstacles
-simultaneously.  We conducted experiments using 4 human subjects who walked
-through the environment 8 times with different configurations of objects, in
-each of these 4 conditions.  We asked whether an agent with modules trained for
-these three sub-tasks could generate paths that matched human's behavior.
-
 \section{Modular Inverse Reinforcement Learning}
@@
 values of taking other actions.
-
-\section{Experiments}
-\label{sec:exp}
-
-We trained the modules first to obtain $Q_i$ before running the inverse
-reinforcement learning algorithm. For each module, the agent only considers the
-closest target and the closest obstacle. For the path module, the path was
-defined as segments connected by waypoints, so only the next waypoint is
-considered. The agent takes the distance and angle to the closest objects as the
-state representation. To make our weights represent the significance of the
-modules, we normalized the sub-MDPs with the unit (positive or negative) reward.
-The reward is 1 for collecting a target, -1 for running into an obstacle, and 1
-for collecting a waypoint of the path. In the implementation, we made the radius
-of waypoints larger than that of other objects, so that the agent will not cling
-too much to the path.
-
-We set some constraints on the learning agent in order to make it walk roughly
-like a human while keeping the action space small.  We observed from the human
-trajectories that humans walk smoothly and do not turn sharply.  Our agent was
-only allowed to do three actions --- going straight ahead (0.3 meter),
-turning left with a small step (30 degrees to the left), and turning right with
-a small step (30 degrees to the right).
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_1.png}
-\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_2.png}
-\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_3.png}
-\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_4.png}
-\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
-\end{subfigure}
-\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
-black lines are trajectories of human subjects, and the green lines are
-trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights for each task are given as (target,
-obstacle, path).}
-
-\label{fig:exp}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact1.png}
-\caption{Path module only. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact2.png}
-\caption{Obstacle + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact3.png}
-\caption{Target + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{contact4.png}
-\caption{Target + Obstacle + Path. }
-\end{subfigure}
-\caption{Number of targets hit and number of obstacles hit of the human subjects 
-and the agent.}
-%COMMENT Shun it would be much better if you used black and green in this figure instead of blue and red since it agree 
-%with your previous figure. Also as a side comment its rather obvious that you could do better by using a path constraint that was not so severe
-\label{fig:stats}
-\end{figure}
-
-The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
-the red circles are obstacles and the blue circles are targets. The gray lines
-are the path. The black lines are trajectories of individual human subjects,
-with each line representing one human trajectory.  Using the weights recovered
-by IRL, the trajectories of the agents are drawn in green. Although individual
-subjects varied in how they avoided obstacles or collected targets, the agent
-can figure out what the humans are doing on aggregate, and perform similarly.
-Figure~\ref{fig:stats} evaluates the performance by showing the number of
-targets hit and number of obstacles hit.  The number of targets hit should be
-large, and the number of obstacles hit should be small, when the corresponding
-module is active.  We can observe that humans still do better than the agent in
-these tasks.  Nevertheless, the agent performance is quite similar to that of
-the human subjects.
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask1.png}
-\caption{Path following only.}
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask2.png}
-\caption{Obstacle + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask3.png}
-\caption{Target + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask4.png}
-\caption{Target + Obstacle + Path. }
-\end{subfigure}
-\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for
-different weights for the four tasks, respectively. The white zones indicate
-higher probabilities. The weights of all three modules sum to 1, so we only show
-the weights on the target and the obstacle modules.
-}
-\label{fig:heatmap}
-\end{figure}
-
-In Figure~\ref{fig:heatmap}, we show the $\log$ of values of
-Equation~\ref{eq:irl} for different weights. The white color represents higher
-probability. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so neither the target module nor the
-obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find that fairly well-defined optima exist for
-these tasks. The optimal weights are also consistent with the experiment
-context.
-
-\section{Conclusions}
-\label{sec:conclude}
-
-We analyzed human behavior using inverse modular reinforcement learning. The
-experimental results show that modular reinforcement learning can explain human
-behavior well, even though the performance of the agent is currently inferior to
-human subjects'.
-
-Note that a weighted sum of Q functions is just one way to combine multiple
-sub-MDPs. Other ways are possible, including, for example, scheduling between
-different modules, with only one active at one time. This is also called {\em
-skill} in the literature \cite{konidaris2009skill}. However, we adopt the
-weighted sum approach because this is more reasonable for human behavior. When a
-human tries to collect targets while avoiding obstacles, these two modules are
-expected to be both active. A scheduling approach may yield frequent oscillation
-between these two modules. Note also that we assume independence between
-modules. However, correlation between modules doesn't impair our analysis in
-this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
-obstacle module tend to be negatively correlated from the shape of the white
-zones. Lastly, weights may be dynamic and different from state to state.
-However, with such an assumption we need to learn a mapping from state to
-weights. In this case, the curse of dimensionality still exists, and inverse
-learning would be difficult.

### 2015-03-28 4-experiments.tex
migrate to nips

--- 2015-02-20 content.tex
@@
-\section{Introduction}
-
-A large body of evidence from human and other primate studies 
-suggests that they use reinforcement signals to learn tasks and reinforcement 
-learning has proved to be a good model for this process.
-However humans are able to learn and carry out very complex tasks involving 
-many different objectives. 
-How can they do this? The likelihood is that they use reinforcement in this 
-context as well, but most formal reinforcement learning models do not scale 
-well with increasing complexity.
-
-One promising possibility is that the complex task can be broken down into 
-sub-tasks that are each learned separately \cite{sprague2003multiple,
-rothkopf2013modular, dietterich2000hierarchical}. 
-For certain task venues, this decomposition allows the behavior in the complex 
-task to be chosen based on the value of a weighted sum of individual sub-task priorities.
-This paper explores the use of such a decomposition to characterize complex
-behavior of subjects in a virtual reality environment. The methodology of
-\cite{rothkopf2013modular} allows the task priorities to be estimated using the
-subjects' behavioral data. Our experimental analysis shows that the modular reinforcement can be a 
-surprisingly good model of behavior, predicting individual subjects' sub-task 
-priorities in a way that explains their behavioral choices.
-
-This paper is organized as follows. Section~\ref{sec:domain} introduces the
-domain and the composite task structure that we collected human data. Section~\ref{sec:rl}
-describes the main algorithm, modular inverse reinforcement learning. We report
-our experiment results in Section~\ref{sec:exp}, and conclude in
-Section~\ref{sec:conclude}.
-
-\section{Multi-objective Domain}
-\label{sec:domain}
-
-\begin{figure}[h]
-\centering
-\includegraphics[height=5cm]{human.jpg}
-\includegraphics[height=6cm]{env.png}
-\caption{(Left) A human subject with a head mounted display (HMD) and trackers
-for the eye, head, and body.  (Right) The environment the human can see through
-the HMD.  The red cubes represent obstacles. The blue balls represent targets.
-There is also a gray path on the ground that the human subject can follow.}
-\label{fig:avatar}
-\end{figure}
-
-Figure~\ref{fig:avatar} shows the experimental domain. The
-human subject was immersed in a virtual reality domain by wearing a binocular head-mounted display.
-The subject's eye, head, and body motion were tracked as he/she walked through a
-virtual environment that was designed to match the dimensions of a standard indoor
-environment. The subject was asked to follow the path, avoid the obstacles (red
-cubes), and collect the targets (blue spheres) by intercepting them. In
-different conditions they were asked to give different importance to particular
-sub-tasks, with sub-tasks being either relevant or irrelevant. Subjects were
-given auditory feedback when running into obstacles, and when intercepting
-targets, depending on their importance in the current condition. Thus this
-domain can be modeled by three modules, 1) following the path, 2) collecting targets, 3)
-avoiding obstacles.  This general paradigm has been used to evaluate modular
-reinforcement learning \cite{rothkopf2013modular} and understand human behavior
-\cite{Tong2014}.
-
-We evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoiding the
-obstacles.  {\bf Task 3}, following the path, while collecting targets. {\bf
-Task 4}, following the path, collecting the targets and avoiding obstacles
-simultaneously.  We conducted experiments using 4 human subjects who walked
-through the environment 8 times with different configurations of objects, in
-each of these 4 conditions.  We asked whether an agent with modules trained for
-these three sub-tasks could generate paths that matched human's behavior.
-
-\section{Modular Inverse Reinforcement Learning}
-\label{sec:rl}
-
-In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
-function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
-The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
-how do we determine the global policy from the modules, or sub-tasks?
-
-%In the
-%literature, work has been done to integrate the decomposed value functions of
-%the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
-%own policies, while the global policy is a weighted sum of those sub-task
-%policies \cite{thomas2012motor}.
-
-%COMMENT: Shun, the modular formulation started with Sprague et al ACM Trans. of Applied Perception 4(2), 1 , 2007 and this should be referenced.
-%The Parr and Koller paper is about factored MDPs which is diofferent than modular MDPs. The Thomas and Barto is a much more recent paper
-%that learns motor MDPs in a very restructed domain.
-In the literature, modular formation is proposed for understanding human
-visuomotor behavior \cite{sprague2007modeling}. A combination of various basic
-modules can generate complex behavior to match human behavior.
-Following \cite{sprague2007modeling, rothkopf2013modular}, we consider how the
-modules can be combined. Here we assume that the global Q function is a weighted
-sum of all $Q_i$, where $Q_i$ is the Q function for i-th module.
-$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
-where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-$s_i$ denotes the decomposition of $s$ in the i-th module.
-
-Different weights can yield different performance. Let $w$ be the vector of
-$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target et al 
-collection, obstacle avoidance, and path following, respectively. An agent with
-$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
-avoid the obstacles and follow the path.
-
-To obtain the weights given the samples of observed human behavior, we use inverse
-reinforcement learning. In reinforcement learning, we derive 
-optimal policies given an MDP. In inverse reinforcement learning, however, aims to
-find out the underlying MDP by observing policies. A common
-way is to use a maximum likelihood method to recover the transition function and
-the reward function \cite{ng2000algorithms}, but this approach is very expensive
-as some kind of learning is used in the innermost search loop. The modular
-approach allows a much more economical approach \cite{rothkopf2013modular},
-since we module weights are the only free parameters. The transition function is
-known, and the reward function is trivially the weighted sum of that of modules.
-More precisely, we use the modular inverse reinforcement learning method in
-\cite{rothkopf2013modular} to maximize the function below.
-\begin{equation}
-\label{eq:irl}
-\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
-\end{equation}
-where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
-$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
-before. $\eta$ is a hyperparameter that determines the consistency of human's
-behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.
-
-The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
-sample, then the Q value of taking that action should be larger compared to Q
-values of taking other actions.
-
 \section{Experiments}
@@
 context.
-
-\section{Conclusions}
-\label{sec:conclude}
-
-We analyzed human behavior using inverse modular reinforcement learning. The
-experimental results show that modular reinforcement learning can explain human
-behavior well, even though the performance of the agent is currently inferior to
-human subjects'.
-
-Note that a weighted sum of Q functions is just one way to combine multiple
-sub-MDPs. Other ways are possible, including, for example, scheduling between
-different modules, with only one active at one time. This is also called {\em
-skill} in the literature \cite{konidaris2009skill}. However, we adopt the
-weighted sum approach because this is more reasonable for human behavior. When a
-human tries to collect targets while avoiding obstacles, these two modules are
-expected to be both active. A scheduling approach may yield frequent oscillation
-between these two modules. Note also that we assume independence between
-modules. However, correlation between modules doesn't impair our analysis in
-this paper. In Figure~\ref{fig:heatmap}, we can tell that the target module and
-obstacle module tend to be negatively correlated from the shape of the white
-zones. Lastly, weights may be dynamic and different from state to state.
-However, with such an assumption we need to learn a mapping from state to
-weights. In this case, the curse of dimensionality still exists, and inverse
-learning would be difficult.

### 2015-03-28 5-conclusion.tex
migrate to nips

--- 2015-02-20 old/paper.tex
@@
-\documentclass[11pt]{article} % For LaTeX2e
-\usepackage{rldmsubmit,palatino}
-\usepackage{graphicx}
-\usepackage{sidecap}
-\usepackage{caption}
-\usepackage{subcaption}
-
-\title{Modular Inverse Reinforcement Learning on Human Motion}
-
-\author{
-Shun Zhang\\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{menie482@cs.utexas.edu} \\
-\And
-Matthew Tong \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{mhtong@gmail.com} \\
-\AND
-Mary Hayhoe \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{hayhoe@utexas.edu} \\
-\And
-Dana Ballard \\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{dana@cs.utexas.edu} \\
-\\
-}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to \LaTeX{} to determine where to break
-% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
-% puts 3 of 4 authors names on the first line, and the last on the second
-% line, try using \AND instead of \And before the third author name.
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-
-\newcommand{\commentd}[1]{{\bf **Dana: #1**}}
-
-\begin{document}
-
-\maketitle
-
-%The abstract should be a maximum of 2000 characters of text, including
-%spaces (no figure is allowed).
-\begin{abstract}
-Reinforcement learning has been seen as a useful model for understanding human
-behavior because of the importance of the neural reward circuitry. However,
-because of the difficulty of scaling up RL models to large state spaces, it has
-been hard to apply these models to complex human behaviors. One potential
-simplification, consistent with observations of natural behavior, is that
-complex tasks can be broken down into independent sub-tasks, or modules . In
-this paper, we use observed human behavior while walking along a path to
-estimate the intrinsic reward values associated with different modular
-sub-tasks. To do this we use a simplified version of Inverse Reinforcement
-Learning to calculate the reward associated with path following, obstacle
-avoidance, and target collection of humans acting in an immersive virtual
-environment. Using the estimated values, a modular RL model can generate
-realistic behavior consistent with human action choices. This provides a way of
-understanding momentary sensorimotor decisions made in complex natural
-environments.
-\end{abstract}
-
-\keywords{
-Reinforcement learning, human motion, inverse learning
-}
-
-\startmain % to start the main 1-4 pages of the submission.
-
-\section{Introduction}
-
-A large body of evidence from human and other primate studies 
-suggests that they use reinforcement signals to learn tasks and reinforcement 
-learning has proved to be a good model for this process.
-However humans are able to learn and carry out very complex tasks involving 
-many different objectives. 
-How can they do this? The likelihood is that they use reinforcement in this 
-context as well, but most formal reinforcement learning models do not scale 
-well with increasing complexity.
-
-One promising possibility is that the complex task can be broken down into 
-sub-tasks that are each learned separately \cite{sprague2003multiple,
-rothkopf2013modular, dietterich2000hierarchical}. 
-For certain task venues, this decomposition allows the behavior in the complex 
-task to be chosen based on the value of a weighted sum of individual sub-task priorities.
-This paper explores the use of such a decomposition to characterize complex
-behavior of subjects in a virtual reality environment. The methodology of
-\cite{rothkopf2013modular} allows the task priorities to be estimated using the
-subjects' behavioral data. Our experimental analysis shows that the modular reinforcement can be a 
-surprisingly good model of behavior, predicting individual subjects' sub-task 
-priorities in a way that explains their behavioral choices.
-
-This paper is organized as follows. Section~\ref{sec:domain} introduces the
-domain and the composite task structure that we collected human data. Section~\ref{sec:rl}
-describes the main algorithm, modular inverse reinforcement learning. We report
-our experiment results in Section~\ref{sec:exp}, and conclude in
-Section~\ref{sec:conclude}.
-
-\section{Multi-objective Domain}
-\label{sec:domain}
-
-\begin{SCfigure}[1.0][h]
-\centering
-\includegraphics[height=5cm]{human.jpg}
-\includegraphics[height=6cm]{env.png}
-\caption{(Left) A human subject with a head mounted display (HMD) and trackers
-for the eye, head, and body.  (Right) The environment the human can see through
-the HMD.  The red cubes represent obstacles. The blue balls represent targets.
-There is also a gray path on the ground that the human subject can follow.}
-\label{fig:avatar}
-\end{SCfigure}
-
-Figure~\ref{fig:avatar} shows the experimental domain. The
-human subject was immersed in a virtual reality domain by wearing a binocular head-mounted display.
-The subject's eye, head, and body motion were tracked as he/she walked through a
-virtual environment that was designed to match the dimensions of a standard indoor
-environment. The subject was asked to follow the path, avoid the obstacles (red
-cubes), and collect the targets (blue spheres) by intercepting them. In
-different conditions they were asked to give different importance to particular
-sub-tasks, with sub-tasks being either relevant or irrelevant. Subjects were
-given auditory feedback when running into obstacles, and when intercepting
-targets, depending on their importance in the current condition. Thus this
-domain can be modeled by three modules, 1) following the path, 2) collecting targets, 3)
-avoiding obstacles.  This general paradigm has been used to evaluate modular
-reinforcement learning \cite{rothkopf2013modular} and understand human behavior
-\cite{Tong2014}.
-
-We evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoiding the
-obstacles.  {\bf Task 3}, following the path, while collecting targets. {\bf
-Task 4}, following the path, collecting the targets and avoiding obstacles
-simultaneously.  We conducted experiments using 4 human subjects who walked
-through the environment 8 times with different configurations of objects, in
-each of these 4 conditions.  We asked whether an agent with modules trained for
-these three sub-tasks could generate paths that matched human's behavior.
-
-\section{Modular Inverse Reinforcement Learning}
-\label{sec:rl}
-
-In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
-function $Q(s, a)$ evaluates the utility of taking action $a$ from state $s$.
-The {\em policy} of a state is the action with the maximum Q value \cite{rl}. So
-how do we determine the global policy from the modules, or sub-tasks?
-
-%In the
-%literature, work has been done to integrate the decomposed value functions of
-%the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
-%own policies, while the global policy is a weighted sum of those sub-task
-%policies \cite{thomas2012motor}.
-\textcolor{red}{COMMENT - DO NOT LEAVE IN Shun, the modular formulation started with Sprague et al ACM Trans. of Applied Perception 4(2), 1 , 2007 and this should be referenced.
-The Parr and Koller paper is about factored MDPs which is diofferent than modular MDPs. The Thomas and Barto is a much more recent paper
-that learns motor MDPs in a very restructed domain.}
-
-Following (Sprague and Ballard 2007, Rothkopf and Ballard 2013), we assume that the global Q function is a weighted sum of all $Q_i$, where
-$Q_i$ is the Q function for i-th module.
-$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
-where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-$s_i$ denotes the decomposition of $s$ in the i-th module.
-
-Different weights can yield different performance. Let $w$ be the vector of
-$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target et al 
-collection, obstacle avoidance, and path following, respectively. An agent with
-$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
-avoid the obstacles and follow the path.
-
-To obtain the weights given the samples of observed human behavior, we use inverse
-reinforcement learning. In reinforcement learning, we derive 
-optimal policies given an MDP. In inverse reinforcement learning, however, aims to
-find out the underlying MDP by observing policies. A common
-way is to use a maximum likelihood method to recover the transition function and
-the reward function \cite{ng2000algorithms}, but this approach is very expensive as some kind of learning
-is used in the innermost search loop. The modular approach allows a much more economical approach (Rothkopf and Ballard), since we module weights
-are the only free parameters. The transition function is known, and the reward function is
-trivially the weighted sum of that of modules. More precisely, we use the modular
-inverse reinforcement learning method in \cite{rothkopf2013modular} to maximize the
-function below.
-\begin{equation}
-\label{eq:irl}
-\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
-\end{equation}
-where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
-$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
-before. $\eta$ is a hyperparameter that determines the consistency of human's
-behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.
-
-The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
-sample, then the Q value of taking that action should be larger compared to Q
-values of taking other actions.
-
-\section{Experiments}
-\label{sec:exp}
-
-We trained the modules first to obtain $Q_i$ before running the inverse
-reinforcement learning algorithm. For each module, the agent only considers the
-closest target and the closest obstacle. For the path module, the path was
-defined as segments connected by waypoints, so only the next waypoint is
-considered. The agent takes the distance and angle to the closest objects as the
-state representation. To make our weights represent the significance of the
-modules, we normalized the sub-MDPs with the unit (positive or negative) reward.
-The reward is 1 for collecting a target, -1 for running into an obstacle, and 1
-for collecting a waypoint of the path. In the implementation, we made the radius
-of waypoints larger than that of other objects, so that the agent will not cling
-too much to the path.
-
-We set some constraints on the learning agent in order to make it walk roughly
-like a human while keeping the action space small.  We observed from the human
-trajectories that humans walk smoothly and do not turn sharply.  Our agent was
-only allowed to do three actions --- going straight ahead, turning left with a
-small step, and turning right with a small step. \textcolor{red}{COMMENT It would be good to say how much in degrees}
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_1.png}
-\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_2.png}
-\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_3.png}
-\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
-\end{subfigure}
-\begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_4.png}
-\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
-\end{subfigure}
-\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
-black lines are trajectories of human subjects, and the green lines are
-trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights for each task are given as (target,
-obstacle, path).}
-
-\label{fig:exp}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{contactStats.png}
-\caption{Number of targets hit and number of obstacles hit of the human subjects 
-and the agent.\textcolor{red}{COMMENT Shun it would be much better if you used black and green in this figure instead of blue and red since it agree 
-with your previous figure. Also as a side comment its rather obvious that you could do better by using a path constraint that was not so severe}}
-\label{fig:stats}
-\end{figure}
-
-The results are shown in Figure~\ref{fig:exp}. As in Figure~\ref{fig:avatar},
-the red circles are obstacles and the blue circles are targets. The gray lines
-are the path. The black lines are trajectories of individual human subjects,
-with each line representing one human trajectory.  Using the weights recovered
-by IRL, the trajectories of the agents are drawn in green. Although individual
-subjects varied in how they avoided obstacles or collected targets, the agent
-can figure out what the humans are doing on aggregate, and perform similarly.
-Figure~\ref{fig:stats} evaluates the performance by showing the number of
-targets hit and number of obstacles hit.  The number of targets hit should be
-large, and the number of obstacles hit should be small, when the corresponding
-module is active.  We can observe that humans still do better than the agent in
-these tasks.  Nevertheless, the agent performance is quite similar to that of
-the human subjects.
-
-\begin{figure}[h]
-\centering
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask1.png}
-\caption{Path following only.}
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask2.png}
-\caption{Obstacle + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask3.png}
-\caption{Target + Path. }
-\end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
-\includegraphics[width=\textwidth]{objValuesTask4.png}
-\caption{Target + Obstacle + Path. }
-\end{subfigure}
-\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for
-different weights for the four tasks, respectively. The white zones indicate
-higher probabilities. The weights of all three modules sum to 1, so we only show
-the weights on the target and the obstacle modules.
-}
-\label{fig:heatmap}
-\end{figure}
-
-In Figure~\ref{fig:heatmap}, we show the $\log$ of values of
-Equation~\ref{eq:irl} for different weights. The white color represents higher
-probability. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so neither the target module nor the
-obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find that fairly well-defined optima exist for
-these tasks. The optimal weights are also consistent with the experiment
-context.
-
 \section{Conclusions}
@@
 learning would be difficult.
-
-\bibliographystyle{plain}
-\bibliography{paper}
-
-\end{document}

### 2015-03-28 paper.tex
update content
migrate to nips

@@
-\documentclass[11pt]{article} % For LaTeX2e
-\usepackage{rldmsubmit,palatino}
+\documentclass{article} % For LaTeX2e
+\usepackage{nips15submit_e,times}
+\usepackage{hyperref}
+\usepackage{url}
 \usepackage{graphicx}
+\usepackage{sidecap}
 \usepackage{caption}
 \usepackage{subcaption}
+%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
 
@@
 
-\author{
-Shun Zhang\\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{menie482@cs.utexas.edu} \\
-\And
-Matthew Tong \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{mhtong@gmail.com} \\
-\AND
-Mary Hayhoe \\
-Center for Perceptual Systems\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{hayhoe@utexas.edu} \\
-\And
-Dana Ballard \\
-Department of Computer Science\\
-University of Texas at Austin\\
-Austin, TX 78712 \\
-\texttt{dana@cs.utexas.edu} \\
-\\
-}
+%\author{
+%}
 
@@
 
-\newcommand{\commentd}[1]{{\bf **Dana: #1**}}
+%\nipsfinalcopy % Uncomment for camera-ready version
 
@@
 
-%The abstract should be a maximum of 2000 characters of text, including
-%spaces (no figure is allowed).
 \begin{abstract}
@@
 
-\keywords{
-Reinforcement learning, human motion, inverse learning
-}
-
-\startmain % to start the main 1-4 pages of the submission.
-
-\include{content}
+\input{1-intro}
+\input{2-domain}
+\input{3-irl}
+\input{4-experiments}
+\input{5-conclusion}
 

### 2015-03-29 3-irl.tex
minor changes in irl

@@
 
+% TODO expand this
 In Reinforcement Learning, when given a state and action pair, $s$ and $a$, the
@@
 
-%In the
-%literature, work has been done to integrate the decomposed value functions of
-%the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
-%own policies, while the global policy is a weighted sum of those sub-task
-%policies \cite{thomas2012motor}.
-
-%COMMENT: Shun, the modular formulation started with Sprague et al ACM Trans. of Applied Perception 4(2), 1 , 2007 and this should be referenced.
-%The Parr and Koller paper is about factored MDPs which is diofferent than modular MDPs. The Thomas and Barto is a much more recent paper
-%that learns motor MDPs in a very restructed domain.
 In the literature, modular formation is proposed for understanding human
@@
 of modules \cite{Rothkopf12Infer}.
-Following \cite{sprague2007modeling, rothkopf2013modular}, we consider how the
+Other work has been done to integrate the decomposed value functions of
+the sub-tasks \cite{koller1999computing}. The sub-tasks can also propose their
+own policies, while the global policy is a weighted sum of those sub-task
+policies \cite{thomas2012motor}.
+
+We now consider how the
 modules can be combined. Here we assume that the global Q function is a weighted
-sum of all $Q_i$, where $Q_i$ is the Q function for i-th module.
+sum of all $Q_i$ \cite{sprague2007modeling, rothkopf2013modular}, where $Q_i$ is
+the Q function for i-th module.
 $$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
@@
 
+% TODO mention discounters here
 To obtain the weights given the samples of observed human behavior, we use inverse
@@
 \cite{rothkopf2013modular} to maximize the function below.
+% TODO more details
 \begin{equation}

### 2015-05-29 present.tex
poster init

--- 2015-02-14 content.tex
@@
-\section{Introduction}
+\documentclass[final]{beamer}
+\usetheme{RJH}
+\usepackage[width=180,height=120,scale=1.5,debug]{beamerposter}
+\usepackage[absolute,overlay]{textpos}
+\setlength{\TPHorizModule}{1cm}
+\setlength{\TPVertModule}{1cm}
 
-A large body of evidence from human and other primate studies 
-suggest that they use reinforcement signals to learn tasks and reinforcement 
-learning has proved to be a good model for this process.
-However humans are able to learn and carry out very complex tasks involving 
-many different objectives. 
-How can they do this? The likelihood is that they use reinforcement in this 
-context as well but most formal reinforcement learning models do not scale 
-well with increasing complexity.
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 
-One promising possibility is that the complex task can be broken down into 
-sub-tasks that are each learned separately \cite{sprague2003multiple,
-rothkopf2013modular, dietterich2000hierarchical}. 
-For certain task venues, this decomposition allows the behavior in the complex 
-task to be specified as a weighted sum of individual sub-task priorities.
-This paper explores the use of such a decomposition to characterize complex
-behavior of subjects in a virtual reality environment. The methodology of
-\cite{rothkopf2013modular} allows the task priorities to be estimated using the
-subjects' behavioral data.
+\title{Modular Inverse Reinforcement Learning on Human Motion}
+\author{Shun Zhang, Matthew Tong, Mary Hayhoe, Dana Ballard\\
+Department of Computer Science, Center for Perceptual Systems\\
+University of Texas at Austin}
+\footer{}
 
-Our experimental analysis shows that the modular reinforcement can be a 
-surprisingly good model of behavior, predicting individual subjects' sub-task 
-priorities in a way that explains their behavioral choices.
+\begin{document}
+\begin{frame}{}
 
-This paper is organized as follows. Section~\ref{sec:domain} introduces the
-domain of the composite task that we collected human data. Section~\ref{sec:rl}
-describes the main algorithm, modular inverse reinforcement learning. We report
-our experiment results in Section~\ref{sec:exp}, and conclude in
-Section~\ref{sec:conclude}.
+% left
+\begin{textblock}{37}(1,8)
+\begin{block}{Introduction}
+blahh
+\end{block}
+\end{textblock}
 
-\section{Multi-objective Domain}
-\label{sec:domain}
-
-\begin{figure}[h!]
+% middle
+\begin{textblock}{37}(40,8)
+\begin{block}{Multi-objective Domain}
+\begin{figure}[h]
 \centering
-\includegraphics[height=5cm]{human.jpg}
-\includegraphics[height=5cm]{env.png}
-\caption{(Left) A human subject with head tracker, virtual reality displayer, and
-body tracker.  (Right) The environment the human can see through the displayer.
-The red cubes represent obstacles. The blue balls represent targets. There is
-also a gray path on the ground that the human subject can follow.}
+\includegraphics[width=0.2\textwidth]{human.jpg}
+\includegraphics[width=0.6\textwidth]{env.png}
+\caption{(Left) A human subject with a head mounted display (HMD) and trackers
+for the eye, head, and body.  (Right) The environment the human can see through
+the HMD.  The red cubes represent obstacles. The blue balls represent targets.
+There is also a gray path on the ground that the human subject can follow.}
 \label{fig:avatar}
 \end{figure}
+\end{block}
 
-Figure~\ref{fig:avatar} shows the domain that we used for our experiment. The
-agent, which is a human subject in this case, is in a virtual reality domain.
-We ask the human subject to have trackers and a virtual reality displayer
-equipped. He can see the environment through the displayer in front of his eyes,
-and he can walk as if he is walking in the virtual domain.
-The agent would be punished if running into obstacles (red cubes), and rewarded
-if collecting targets (blue balls). There is a path on the ground that he can
-follow.
-Naturally, this domain has three sub-tasks, 1) following the path, 2)
-collecting targets, 3) avoiding obstacles.
-This is an experiment used in the literature to evaluate modular reinforcement
-learning \cite{rothkopf2013modular} and understand human behavior \cite{Tong2014}.
+\begin{block}{Experiments}
+\begin{figure}[h]
+\centering
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{task_1.png}
+\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
+\end{subfigure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{task_2.png}
+\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
+\end{subfigure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{task_3.png}
+\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
+\end{subfigure}
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=\textwidth]{task_4.png}
+\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
+\end{subfigure}
+\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
+black lines are trajectories of human subjects, and the green lines are
+trajectories of the learning agent by using the optimum weights, $w$, derived
+from modular inverse reinforcement learning. Weights for each task are given as (target,
+obstacle, path).}
+\label{fig:exp}
+\end{figure}
 
-We will evaluate four different tasks. {\bf Task 1}, following the path only, and
-ignoring other objects. {\bf Task 2}, following the path, while avoiding the obstacles.
-{\bf Task 3}, following the path, while collecting targets when possible. {\bf Task 4},
-following the path, collecting the targets and avoiding obstacles
-simultaneously.
-We conducted experiments to ask human subjects to accomplish these tasks and
-recorded their trajectories. The human data are collected by the Center for
-Perceptual Systems at University of Texas at Austin.
+\end{block}
+\end{textblock}
 
-If the agent observes the distance and angle to an object, he is expected to
-know the optimal action to avoid or pursue it. This decision is also Markov.
-Therefore, the question would be, if we have trained modules for these three
-sub-tasks, could they be integrated to match human's behavior?
-
-\section{Modular Inverse Reinforcement Learning}
-\label{sec:rl}
-
-Recall that for a state, action
-pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
-state $s$. The {\em policy} of a state is the action with the maximum Q
-value \cite{rl}. So how do we determine the global policy from the modules, or
-sub-tasks? In the literature, work has been done to integrate the decomposed
-value functions of the sub-tasks \cite{koller1999computing}. The sub-tasks can
-also propose their own policies, while the global policy is a weighted sum of
-those sub-task policies \cite{thomas2012motor}.
-
-Here, we assume that the global Q function is a weighted sum of all $Q_i$, where
-$Q_i$ is the Q function for i-th module.
-$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
-where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-$s_i$ denotes the decomposition of $s$ in the i-th module.
-
-Different weights can yield different performance. Let $w$ be the vector of
-$(w_1, w_2, w_3)$, where $w_1, w_2, w_3$ are weights for the sub-tasks of target
-collection, obstacle avoidance, and path following, respectively. An agent with
-$w = (1, 0, 0)$ only collects targets, and one with $w = (0, 0.5, 0.5)$ may
-avoid the obstacles and follow the path.
-
-To obtain the weights given the samples, we need to use the inverse
-reinforcement learning technique. In reinforcement learning, we derive 
-optimal policies given an MDP. In inverse reinforcement learning, however, we
-find out the underlying MDP by observing policies. In the literature, a common
-way is to use a maximum likelihood method to recover the transition function and
-the reward function \cite{ng2000algorithms}. Here, however, we only need to find
-out the weights. The transition function is known, and the reward function is
-trivially the weighted sum of that of modules. Concretely, we use the modular
-inverse reinforcement learning method \cite{rothkopf2013modular} to maximize the
-function below.
-\begin{equation}
-\label{eq:irl}
-\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
-\end{equation}
-where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
-$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
-before. $\eta$ is a hyperparameter that determines the consistency of human's
-behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.
-
-The intuition of Equation~\ref{eq:irl} is that if an action is observed from the
-sample, then the Q value of taking that action should be larger compared to Q
-values of taking other actions.
-
-\section{Experiments}
-\label{sec:exp}
-
-We trained the modules first to obtain $Q_i$ before running the inverse reinforcement
-learning algorithm. For each module, the agent only considers the closest target
-and the closest obstacle. For the path module, the path is defined as segments
-connected by waypoints, so the next waypoint is considered. The agent takes the
-distance and angle to the closest objects as the state representation.
-
-To make our weights represent the significance of the modules, we
-normalize the sub-MDPs with the unit (positive or negative) reward. The reward
-is 1 for collecting a target, -1 for running into an obstacle, and 1 for
-collecting a waypoint of the path. In the implementation, we make the radius of
-waypoints larger than that of other objects, so that the agent will not cling
-too much to the path.
-
-We make some constraints on our learning agent to make it walk like a human.
-We can find from the human trajectories that humans walk smoothly and do not turn
-sharply.  Our agent is only allowed to do three actions --- going straight ahead,
-turning left with a small step, and turning right with a small step.
-
+% right
+\begin{textblock}{37}(79,8)
+\begin{block}{Experiments}
 \begin{figure}[h]
@@
 \begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_1.png}
-\caption{Path module only,\\$w = (0.039, 0.0, 0.960)$. }
+\includegraphics[width=\textwidth]{contact1.png}
+\caption{Path module only. }
 \end{subfigure}
 \begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_2.png}
-\caption{Obstacle + Path,\\$w = (0.081, 0.264, 0.654)$. }
+\includegraphics[width=\textwidth]{contact2.png}
+\caption{Obstacle + Path. }
 \end{subfigure}
 \begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_3.png}
-\caption{Target + Path, \\$w = (0.254, 0.089, 0.655)$. }
+\includegraphics[width=\textwidth]{contact3.png}
+\caption{Target + Path. }
 \end{subfigure}
 \begin{subfigure}[b]{0.24\textwidth}
-\includegraphics[width=\textwidth]{task_4.png}
-\caption{Target + Obstacle + Path, \\$w = (0.215, 0.414, 0.369)$. }
+\includegraphics[width=\textwidth]{contact4.png}
+\caption{All modules. }
 \end{subfigure}
-\caption{The trajectories of humans and the agent in the four tasks. The
-black lines are trajectories of human subjects, and the green lines are
-trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights are shown in the target,
-obstacle, path order.}
-
-\label{fig:exp}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.5\textwidth]{contactStats.png}
 \caption{Number of targets hit and number of obstacles hit of the human subjects 
@@
 
-We report the results in Figure~\ref{fig:exp}. Same as Figure~\ref{fig:avatar},
-the red circles are obstacles. The blue circles are targets. The gray lines are
-the path. The black lines are trajectories of human --- each line represents one
-human trajectory.
-Using the weights derived from the algorithm, the trajectories of the agents are
-drawn in green color.
-
-Although humans don't agree with each other on how to avoid obstacles or collect
-targets, the agent can figure out what the humans are doing, and perform a
-similar behavior. Figure~\ref{fig:stats} evaluates the
-performance by showing the number of targets hit and number of obstacles hit.
-The number of targets hit should be large, and the number of obstacles hit
-should be small, when the corresponding module is active.
-We can observe that humans still do better than the agent in these tasks.
-Nevertheless, the agent has a similar performance compared to human subjects.
-
 \begin{figure}[h]
 \centering
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.24\textwidth}
 \includegraphics[width=\textwidth]{objValuesTask1.png}
-\caption{Path following only.}
+\caption{Path following.}
 \end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.24\textwidth}
 \includegraphics[width=\textwidth]{objValuesTask2.png}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.24\textwidth}
 \includegraphics[width=\textwidth]{objValuesTask3.png}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.24\textwidth}
 \includegraphics[width=\textwidth]{objValuesTask4.png}
-\caption{Target + Obstacle + Path. }
+\caption{All modules. }
 \end{subfigure}
@@
 \end{figure}
+\end{block}
+\end{textblock}
 
-In Figure~\ref{fig:heatmap}, we show the $\log$ of values of
-Equation~\ref{eq:irl} for different weights. The white color represents higher
-probability. We can observe the centroids of white zones move for different
-tasks. It stays at the origin in Task 1, so none of the target module and
-obstacle module is active. It moves away from the origin when a module is
-active.  From the heatmaps, we find the optimums exist for these tasks. The
-optimal weights are also consistent with the experiment context.
-
-\section{Conclusion and Future Work}
-\label{sec:conclude}
-
-We interpreted human behavior using inverse module reinforcement learning in
-this paper. The experimental results show that modular reinforcement learning
-can well explain human behavior, even though the performance of the agent is
-inferior to human subjects'.
-
-There are also some observations potential for the future work. First, 
-weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
-propose other ways including, for example, scheduling between different modules,
-with only one active at one time. This is also called {\em skill} in the literature
-\cite{konidaris2009skill}. However, we adopt the weighted sum approach
-because this is more reasonable for human behavior. When a human tries to collect
-targets while avoiding obstacles, these two modules are expected to be both
-active. A scheduling approach may yield frequent oscillation between these two
-modules.
-
-Second, we also assume independency between modules. Correlation between
-modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
-we can tell that the target module and obstacle module tend to be negatively
-correlated from the shape of the white zones.
-
-Lastly, weights may be dynamic and different from state to state. However, with
-such assumption, we need to learn a mapping from state to weights. In this case,
-the curse of dimensionality still exists, and inverse learning would be
-difficult.
+\end{frame}
+\end{document}

### 2015-05-29 present.tex
more fleshed out, except intro and irl

@@
 \begin{block}{Introduction}
-blahh
+\begin{itemize}
+\item
+One promising possibility is that the complex task can be broken down into 
+sub-tasks that are each learned separately.
+\end{itemize}
+\end{block}
+
+\begin{block}{Modular Inverse Reinforcement Learning}
+\begin{itemize}
+\item {\bf Reinforcement Learning.}
+\item {\bf Modular Renforcement Learning.}
+We assume that the global Q function is a weighted
+sum of all $Q_i$, where $Q_i$ is the Q function for i-th module.
+$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
+where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
+$s_i$ denotes the decomposition of $s$ in the i-th module.
+\item {\bf Modular Inverse Reinforcement Learning.}
+We maximize the objective function
+$$\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)},
+b)}}$$
+where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
+$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
+before. $\eta$ is a hyperparameter that determines the consistency of human's
+behavior.
+\end{itemize}
 \end{block}
@@
 \end{block}
+
+\begin{block}{Conclusion}
+\begin{itemize}
+\item
+We analyzed human behavior using inverse modular reinforcement learning.
+\item
+The experimental results show that modular reinforcement learning can explain
+human behavior well, even though the performance of the agent is currently
+inferior to human subjects'.
+\end{itemize}
+\end{block}
+
+\begin{block}{Following Work}
+\begin{itemize}
+\item
+Learning weights (or rewards) and discounters of sub-MDPs simutanously.
+\item
+Testing in gridworld domains, with hundreds of sub-MDPs. We compared modular IRL
+and Bayesian IRL in this condition.
+\item
+Evaluation by angular differences in policies, and likelihood of trajectories.
+This is compared with other baseline agents.
+\end{itemize}
+
+{\bf Watch our follow-up paper in a future conference!}
+\end{block}
+
+\begin{block}{Acknowledgement}
+\begin{itemize}
+\item Funding.
+\end{itemize}
+\end{block}
+
 \end{textblock}

### 2015-05-29 present.tex
poster ok except mdp decomp and visual alignments

@@
 \author{Shun Zhang, Matthew Tong, Mary Hayhoe, Dana Ballard\\
-Department of Computer Science, Center for Perceptual Systems\\
+Department of Computer Science, Center of Perceptual Systems\\
 University of Texas at Austin}
@@
 \item
+Humans are able to learn and carry out very complex tasks involving 
+many different objectives, while most reinforcement learning algorithms suffer
+from the curse of dimensionality.
+\item
 One promising possibility is that the complex task can be broken down into 
-sub-tasks that are each learned separately.
+{\bf sub-tasks} that are each learned separately.
+\item
+This {\bf decomposition} allows the behavior in the complex task to be chosen based on
+the value of a weighted sum of individual sub-tasks.
+\item
+Our experimental analysis shows that the modular reinforcement can be a 
+good model of predicting {\bf human subjects'} sub-task priorities in a way that
+explains their behavioral choices.
+\item
+We use the {\bf Modular Inverse Reinforcement Learning} approach to analyze
+human subjects' behaviors.
 \end{itemize}
@@
 \begin{itemize}
-\item {\bf Reinforcement Learning.}
-\item {\bf Modular Renforcement Learning.}
+\item {\bf MDP decomposition.}
+\item {\bf Modular Reinforcement Learning.}
 We assume that the global Q function is a weighted
@@
 \includegraphics[width=\textwidth]{contact1.png}
-\caption{Path module only. }
+\caption{Path module only.}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{contact2.png}
-\caption{Obstacle + Path. }
+\caption{Obstacle + Path.}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{contact3.png}
-\caption{Target + Path. }
+\caption{Target + Path.}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{contact4.png}
-\caption{All modules. }
+\caption{All modules.}
 \end{subfigure}
@@
 \end{subfigure}
-\caption{Heatmaps of the $\log$ of the values of Equation~\ref{eq:irl} for
+\caption{Heatmaps of the $\log$ of the likelihood for
 different weights for the four tasks, respectively. The white zones indicate
@@
 \item
-Learning weights (or rewards) and discounters of sub-MDPs simutanously.
+Learning weights (or rewards) and discounters of sub-MDPs simultaneously.
 \item
-Testing in gridworld domains, with hundreds of sub-MDPs. We compared modular IRL
-and Bayesian IRL in this condition.
+Testing in gridworld domains, with hundreds of sub-MDPs. We compared Modular IRL
+with Bayesian IRL in this condition.
 \item
@@
 
-\begin{block}{Acknowledgement}
+\begin{block}{Acknowledgment}
 \begin{itemize}
 \item Funding.
+\item Ruohan Zhang, on his separate work on modular reinforcement learning.
 \end{itemize}

### 2015-05-30 present.tex
poster fleshed out

@@
 \author{Shun Zhang, Matthew Tong, Mary Hayhoe, Dana Ballard\\
-Department of Computer Science, Center of Perceptual Systems\\
+Department of Computer Science, Center for Perceptual Systems\\
 University of Texas at Austin}
@@
 % left
-\begin{textblock}{37}(1,8)
+\begin{textblock}{35}(1,8)
 \begin{block}{Introduction}
@@
 \begin{itemize}
-\item {\bf MDP decomposition.}
-\item {\bf Modular Reinforcement Learning.}
-We assume that the global Q function is a weighted
-sum of all $Q_i$, where $Q_i$ is the Q function for i-th module.
+\item {\bf Factored MDP.}
+Define $S = S_1 \times S_2 \times \cdots \times S_m$, where $S_i$ is a state
+component. We expect the state components are less correlated. The transition
+function can be defined as a Dynamic Bayesian Network between two time frames of
+state components.
+\item {\bf Q-Decomposition.}
+We further decompose the Q function.
+Assume that the global Q function is a weighted sum of all $Q_i$, where $Q_i$ is
+the Q function for $S_i$, the $i$-th state component.
 $$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
-where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
-$s_i$ denotes the decomposition of $s$ in the i-th module.
+where $w_i$ is a weight scalar. $w_i \geq 0, \sum_i w_i = 1$.
+Now the original MDP is {\bf modularized} into sub-MDPs.
+$$MDP_i = <S_i, A, T, R_i, \gamma>$$
 \item {\bf Modular Inverse Reinforcement Learning.}
-We maximize the objective function
+The objective is to recover the weights given observed samples, $(s^{(i)},
+a^{(i)})$, and sub-MDPs.  We maximize the likelihood of observing such samples,
 $$\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)},
@@
 % middle
-\begin{textblock}{37}(40,8)
+\begin{textblock}{37}(38,8)
 \begin{block}{Multi-objective Domain}
@@
 % right
-\begin{textblock}{37}(79,8)
+\begin{textblock}{39}(77,8)
 \begin{block}{Experiments}
@@
 \includegraphics[width=\textwidth]{objValuesTask1.png}
-\caption{Path following.}
+\caption{Path module only.}
 \end{subfigure}
@@
 different weights for the four tasks, respectively. The white zones indicate
-higher probabilities. The weights of all three modules sum to 1, so we only show
-the weights on the target and the obstacle modules.
+higher probabilities. The vertical axis is the weight of the target module. The
+horizontal axis is the weight of the obstacle module.
 }
@@
 
-{\bf Watch our follow-up paper in a future conference!}
+{\bf More details in our paper to appear in a future conference!}
 \end{block}
@@
 \begin{itemize}
-\item Funding.
-\item Ruohan Zhang, on his separate work on modular reinforcement learning.
+\item Funding?
 \end{itemize}

### 2015-06-02 present.tex
funding added
resize the poster

@@
 \usetheme{RJH}
-\usepackage[width=180,height=120,scale=1.5,debug]{beamerposter}
+\usepackage[size=custom,width=240,height=180,scale=3.3,debug]{beamerposter}
 \usepackage[absolute,overlay]{textpos}
@@
 % left
-\begin{textblock}{35}(1,8)
+\begin{textblock}{73}(2,18)
 \begin{block}{Introduction}
@@
 Define $S = S_1 \times S_2 \times \cdots \times S_m$, where $S_i$ is a state
-component. We expect the state components are less correlated. The transition
-function can be defined as a Dynamic Bayesian Network between two time frames of
-state components.
+component. We expect the state components are less correlated.
 \item {\bf Q-Decomposition.}
@@
 Now the original MDP is {\bf modularized} into sub-MDPs.
-$$MDP_i = <S_i, A, T, R_i, \gamma>$$
 \item {\bf Modular Inverse Reinforcement Learning.}
@@
 % middle
-\begin{textblock}{37}(38,8)
+\begin{textblock}{69}(79,18)
 \begin{block}{Multi-objective Domain}
@@
 \centering
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=\textwidth]{task_1.png}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=\textwidth]{task_2.png}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=\textwidth]{task_3.png}
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{0.47\textwidth}
 \includegraphics[width=\textwidth]{task_4.png}
@@
 \end{subfigure}
-\caption{The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
+\caption{Weights for each task are given as (target,
+obstacle, path).\\
+The trajectories of humans and the agent in the four tasks. Targets are blue and obstacles are red. The
 black lines are trajectories of human subjects, and the green lines are
 trajectories of the learning agent by using the optimum weights, $w$, derived
-from modular inverse reinforcement learning. Weights for each task are given as (target,
-obstacle, path).}
+from modular inverse reinforcement learning. }
 \label{fig:exp}
@@
 % right
-\begin{textblock}{39}(77,8)
+\begin{textblock}{86}(152,18)
 \begin{block}{Experiments}
@@
 
-\begin{block}{Conclusion}
+\begin{block}{Discussion and Conclusion}
 \begin{itemize}
@@
 inferior to human subjects'.
+\item
+The weighted sum of Q functions is just one way to combine multiple
+sub-MDPs. Other ways, for example, scheduling between different modules, are
+also possible.
 \end{itemize}
@@
 \begin{itemize}
-\item Funding?
+\item This work is supported by NIH EY05729.
 \end{itemize}
